### 2023-01-30 outputs/paper/paper.qmd
cleaned 2012 data, simulated data

---
title: "My title"
subtitle: "My subtitle if needed"
author: "Chloe Thierstein"
thanks: "Code and data are available at: https://github.com/cthierst/death_registry_analysis.git."
date: "24 January 2023"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
bibliography: references.bib
---

```{r}
#| include: FALSE

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). Quarto automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

# Data

Our data is of penguins (@fig-bills).

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (@fig-billssssss). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)

```{r}
#| label: fig-billssssss
#| fig-cap: More bills of penguins
#| echo: false
#| fig-width: 8
#| fig-height: 4

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it.

# Model

$$
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
$$ {#eq-bayes}

@eq-bayes seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References

### 2023-01-30 02-data_cleaning.R

@@ -1,0 +1,49 @@
+#### Preamble ####
+# Purpose: Clean the place of death data downloaded from Open Data Toronto's statistics on death registration. Link: https://open.toronto.ca/dataset/death-registry-statistics/
+# Author: Chloe Thierstein
+# Email: chloe.thierstein@utoronto.ca
+# Date: 23 January 2023
+# Prerequisites: Need to know where to find the City of Toronto death registry data
+
+#### workspace set-up ####
+#### install.packages(tidyverse) #####
+#### install.packages(janitor) #####
+
+library(tidyverse)
+library(janitor)
+
+#### Clean Names ####
+cleaned_place_of_death_data_2012 <-
+  clean_names(raw_death_registry_data)
+
+head(cleaned_place_of_death_data_2012)
+
+#### Simplifying Date #### 
+cleaned_place_of_death_data_2012 <-
+  cleaned_place_of_death_data_2012 |>
+  separate(
+    col = time_period,
+    into = c("Year", "Month"),
+    sep = "-"
+  ) |>
+  select(-Month) #remove month column as unnecessary
+
+#### Selecting Relevant Rows ####
+cleaned_place_of_death_data_2012 <- cleaned_place_of_death_data_2012[c(82:155),] #only keeping rows for 2012
+
+
+#### Selecting Relevant Columns####
+cleaned_place_of_death_data_2012 <-
+  cleaned_place_of_death_data_2012 |>
+  select(
+    death_licenses,
+    place_of_death
+  )
+
+
+
+
+
+
+
+

### 2023-01-30 scripts/00-simulation.R

@@ -1,0 +1,55 @@
+#### Preamble ####
+# Purpose: Simulate data from Open Data Toronto's statistics on death registration. Link: https://open.toronto.ca/dataset/death-registry-statistics/
+# include columns ID, Death Licenses, Civic Centre, Place of Death, Month (as placeholder for date)
+# Author: Chloe Thierstein
+# Email: chloe.thierstein@mail.utoronto.ca
+# Date: 23 January 2023
+# Prerequisites: Need to know where to find the City of Toronto death registry data
+
+
+
+library(tibble)
+
+
+simulated_data <-
+  tibble(
+    "ID" = 1:841,
+    "Death Licenses" = runif(n = 841, min = 1, max = 1000),
+    "Civic Centre" = sample(
+      x = c(
+        "Etobicoke",
+        "North York",
+        "Scarborough",
+        "Toronto"
+     ),
+     size = 841,
+     replace = TRUE
+    ),
+    "Place of Death" = sample(
+      x = c(
+        "Toronto",
+        "Outside of City Limits"
+      ),
+      size = 841,
+      replace = TRUE
+    ),
+    "Month" = sample(
+      x = c(
+        "January",
+        "February",
+        "March",
+        "April",
+        "May",
+        "June",
+        "July",
+        "August",
+        "September",
+        "October",
+        "November",
+        "December"
+      ),
+      size = 841,
+      replace = TRUE
+  )
+)
+simulated_data

### 2023-01-30 scripts/01-data_cleaning.R

@@ -1,0 +1,86 @@
+#### Preamble ####
+# Purpose: Read in data from Open Data Toronto and Clean the civic centre data downloaded from Open Data Toronto's statistics on death registration. Link: https://open.toronto.ca/dataset/death-registry-statistics/
+# Author: Chloe Thierstein
+# Email: chloe.thierstein@utoronto.ca
+# Date: 23 January 2023
+# Prerequisites: Need to know where to find the City of Toronto death registry data
+
+#### workspace set-up ####
+#### install.packages(tidyverse) #####
+#### install.packages(janitor) #####
+
+library(tidyverse)
+library(janitor)
+
+#### Read in Data ####
+raw_death_registry_data <-
+  read_csv(
+    file = 
+      "https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/cba07a90-984b-42d2-9131-701c8c2a9788/resource/cc418325-b3d0-4afd-ae64-ccb9b71c549c/download/Death%20Registry%20Statistics%20Data.csv",
+    show_col_types = FALSE,
+    skip = 0
+  )
+
+write_csv(
+  x = raw_death_registry_data,
+  file = "death_registry.csv"
+)
+
+
+#### Basic Cleaning ####
+raw_death_registry_data <-
+  read_csv(
+    file = "death_registry.csv",
+    show_col_types = FALSE
+  )
+
+
+#### Clean Names ####
+cleaned_civic_centre_data_2012 <-
+  clean_names(raw_death_registry_data)
+
+head(cleaned_civic_centre_data_2012)
+
+#### Simplifying Date #### 
+cleaned_civic_centre_data_2012 <-
+  cleaned_civic_centre_data_2012 |>
+  separate(
+    col = time_period,
+    into = c("Year", "Month"),
+    sep = "-"
+  ) |>
+  select(-Month) #remove month column as unnecessary
+
+
+#### Selecting Relevant Rows ####
+cleaned_civic_centre_data_2012 <- cleaned_civic_centre_data_2012[c(82:155),] #only keeping rows for 2012
+
+#### Recoding Civic Centres to be more Meaningful ####
+cleaned_civic_centre_data_2012 <- cleaned_civic_centre_data_2012 |> 
+  mutate( 
+    civic_centre = 
+      recode( 
+        civic_centre, 
+        "ET" = "Etobicoke", 
+        "NY" = "North York", 
[…]

### 2023-01-31 outputs/paper/paper.qmd
Cleaned dataset and data from removed 2021-2022. Began building graph code for paper reorganized files created death registry csv file

@@
 ---
-title: "My title"
+title: "Death Registry Analysis"
 subtitle: "My subtitle if needed"
@@
 ```{r}
-#| include: FALSE
+#| include: false
 
-knitr::opts_chunk$set(echo = TRUE)
+#### Workspace Set-up ####
 
 library(tidyverse)
-library(palmerpenguins)
+library(opendatatoronto)
+library(janitor)
+library(ggplot2)
+library(dplyr)
+library(knitr)
 ```
@@
 # Introduction
+
+{two or three paragraphs of content. and there should be an additional final paragraph that sets out the remainder of the paper}
 
@@
 
+"This dataset includes information relating to the registration of deaths documented by Registry Services staff located in four of the civic centres (Scarborough, North York, Toronto and Etobicoke).
+Registration of death data is entered into the Registry Services Tracking System (RSTS), from which aggregate statistical information is generated to create the dataset.
+The dataset resides in an Oracle database in the City's environment.
+
+The dataset is created in support of the Vital Statistics Act, which is a Provincial legislation.
+The dataset also supports the City's operational requirements and business functions." [@citeopendatatoronto]
+
+"\_id = unique row identifier for Open Data database
+
+civic_centre = civic centre code
+
+death licenses = number of deaths registered in the month
+
+place_of_death = place where death occurred
+
+time_period = month death registered" [@citeopendatatoronto]
+
+\-\--
+
+In this paper I will be analyzing death registry data from four civic centers in the greater Toronto area, Etobicoke, North York, Scarborough, and Toronto.
+First, I will compare rates of death registration from these civic centers over the period of 2011-2022.
+Next, I will compare rates of death registration against place of death in order to discuss the rate of deaths inside and outside the Toronto city limits.
+Finally, I will discuss trends in death registrations from each civic center.
+My analysis will be performed using R [@citeR], tidyverse [@citetidyverse], janitor [@citejanitor], and dplyr [@citedplyr] packages.
+All figures in my paper have been created using ggplot2 [@citeggplot2] and tables have been created with \_\_\_\_\_\_
+
 # Data
 
-Our data is of penguins (@fig-bills).
+data section should thoroughly and precisely discuss the source of the data and the bias this brings (ethical, statistical, and otherwise).
+comprehensively describe and summarize the data using text, graphs, and tables must be made iwth knitr or gt.
+graphs must show the actual data, or as close to it as possible, not summary statistics.
+(Graphs and tables should be cross-references in the text e.g., 'Table 1 shows...')
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| include: FALSE
+
+#### Read in Data ####
+raw_death_registry_data <-
+  read_csv(
+    file = 
+      "https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/cba07a90-984b-42d2-9131-701c8c2a9788/resource/cc418325-b3d0-4afd-ae64-ccb9b71c549c/download/Death%20Registry%20Statistics%20Data.csv",
+    show_col_types = FALSE,
+    skip = 0
+  )
+
+write_csv(
+  x = raw_death_registry_data,
+  file = "death_registry.csv"
+)
+
+```
+
+```{r}
+#| include: false
+#### Clean Names ####
+cleaned_death_registry_data <-
+  clean_names(raw_death_registry_data)
+
+head(cleaned_death_registry_data)
+
+#### Simplifying Date #### 
+cleaned_death_registry_data <-
+  cleaned_death_registry_data |>
+  separate(
+    col = time_period,
+    into = c("year", "month"),
+    sep = "-"
+  ) |>
+  select(civic_centre, death_licenses, place_of_death, year, month)
+
+#### Removing 2021-2022 rows ####
+cleaned_death_registry_data <- cleaned_death_registry_data[-c(746:841), ]
+
+#### Recoding Months to be more Meaningful ####
+cleaned_death_registry_data <- cleaned_death_registry_data |> 
+  mutate( 
+    month = 
+      recode( 
+        month, 
+        "01" = "Jan", 
+        "02" = "Feb", 
+        "03" = "Mar",
+        "04" = "Apr",
+        "05" = "May",
+        "06" = "Jun",
+        "07" = "Jul",
+        "08" = "Aug",
+        "09" = "Sep",
+        "10" = "Oct",
+        "11" = "Nov",
+        "12" = "Dec"
+      ) 
+  )
+
+#### Recoding Civic Centres to be more Meaningful ####
+cleaned_death_registry_data <- cleaned_death_registry_data |> 
+  mutate( 
+    civic_centre = 
+      recode( 
+        civic_centre, 
+        "ET" = "Etobicoke", 
+        "NY" = "North York", 
+        "SC" = "Scarborough", 
+      ) 
+  )
+
+cleaned_death_registry_data <- cleaned_death_registry_data |>
+  filter(!grepl('TO', civic_centre))
+
+#### Ordering Months #####
+cleaned_death_registry_data$month <- factor(cleaned_death_registry_data$month, levels=c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+```
+
+```{r}
 #| echo: false
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
+#### Creating Graph ####
+ggplot(data=cleaned_death_registry_data, aes(x = month, y = death_licenses, fill = month)) +
   theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+  geom_bar(stat = "identity") +
+  labs(
+    x = "Month",
+    y = "Death Licenses"
+  ) +
+  theme(legend.position = "none") +
+  scale_fill_brewer(palette="Set3")
+
+```
+
+```{r}
+#| echo: false
+
+#### Creating Graph ####
+ggplot(data=cleaned_death_registry_data, aes(x = month, y = death_licenses, fill = year)) +
+  theme_minimal() +
+  geom_bar(stat = "identity") +
+  labs(
+    x = "Month",
+    y = "Death Licenses"
+  ) +
+  scale_fill_brewer(palette="Set3")
+
+```
+
+```{r}
+#| echo: false
+
+#### Creating Graph ####
+ggplot(data=cleaned_death_registry_data, aes(x = year, y = death_licenses, fill = month)) +
+  theme_minimal() +
+  geom_bar(stat = "identity") +
+  labs(
+    x = "year",
+    y = "Death Licenses"
+  ) +
+  scale_fill_brewer(palette="Set3")
+```
+
+```{r}
+#| echo: false
+
+#### Creating Graph ####
+ggplot(data=cleaned_death_registry_data, aes(x = year, y = death_licenses, fill = year)) +
+  geom_bar(stat = "identity") +
+  theme_minimal() + 
+  labs(
+    x = "Year",
+    y = "Death Licenses"
+  ) +
+  theme(legend.position = "none")+
+  scale_fill_brewer(palette="Set3")
+
+```
+
+```{r}
+#| echo: false
+
+options(scipen=999)
+#### Creating Graph ####
+ggplot(data=cleaned_death_registry_data, aes(x = place_of_death, y = death_licenses, fill = place_of_death)) +
+  geom_bar(stat = "identity") +
+  theme_minimal() +
+  labs(
+    x = "Place of Death",
+    y = "Death Licenses"
+  ) +
+  theme(legend.position = "none")+
+  scale_fill_brewer(palette="Set2")
+```
+
+```{r}
+#| echo: false
+#### Creating Graph ####
+ggplot(data=cleaned_death_registry_data, aes(x = civic_centre, y = death_licenses, fill = civic_centre)) +
+  geom_bar(stat = "identity") +
+  theme_minimal() + 
+  labs(
+    x = "Civic Centre",
+    y = "Death Licenses"
+  )+
+  scale_fill_brewer(palette="Set2")+
+  theme(legend.position = "none")
+```
+
+```{r}
+#| echo: false
+#### Creating Graph ####
+ggplot(data=cleaned_death_registry_data, aes(x = year, y = death_licenses, fill = civic_centre)) +
+  geom_bar(stat = "identity") +
+  theme_minimal() + 
+  labs(
+    x = "Year",
+    y = "Death Licenses"
+  )+
+  scale_fill_brewer(palette="Set2")
+```
+
+```{r}
+#| echo: false
+#### Creating Graph ####
+ggplot(data=cleaned_death_registry_data, aes(x = month, y = death_licenses, fill = civic_centre)) +
+  geom_bar(stat = "identity") +
+  theme_minimal() + 
+  labs(
+    x = "Month",
+    y = "Death Licenses"
+  ) +
+  scale_fill_brewer(palette="Set2")
 ```
@@
 
-Also bills and their average (@fig-billssssss).
+Also bills and their average .
 (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)
 
-```{r}
-#| label: fig-billssssss
-#| fig-cap: More bills of penguins
-#| echo: false
-#| fig-width: 8
-#| fig-height: 4
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = FALSE) +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
 Talk way more about it.
-
-# Model
-
-$$
-Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
-$$ {#eq-bayes}
-
-@eq-bayes seems useful, eh?
-
-Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR].
-We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-# Discussion
-
-## First discussion point
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Additional details
 
@@
 # References
+
+references should be added using BibTeX. be sure to reference R and any R packages you use, as well as the dataset. strong submissions will draw on related literature and reference those.
+
+\-\-\--

### 2023-01-31 01-data_cleaning.R

@@ -1,0 +1,131 @@
+
+#### Preamble ####
+# Purpose: Read in data from Open Data Toronto and clean the civic centre data from 2012 downloaded from Open Data Toronto's statistics on death registration. Link: https://open.toronto.ca/dataset/death-registry-statistics/
+# Author: Chloe Thierstein
+# Email: chloe.thierstein@utoronto.ca
+# Date: 23 January 2023
+# Prerequisites: Need to know where to find the City of Toronto death registry data
+
+#### workspace set-up ####
+#### install.packages(tidyverse) ####
+#### install.packages(janitor) ####
+#### install.packages("knitr") ####
+
+install.packages("RColorBrewer")
+library("RColorBrewer")
+
+library(knitr)
+library(tidyverse)
+library(janitor)
+
+
+
+
+#### Clean Names ####
+cleaned_death_registry_data <-
+  clean_names(raw_death_registry_data)
+
+head(cleaned_death_registry_data)
+
+
+#### Simplifying Date #### 
+cleaned_death_registry_data <-
+  cleaned_death_registry_data |>
+  separate(
+    col = time_period,
+    into = c("year", "month"),
+    sep = "-"
+  ) |>
+  select(civic_centre, death_licenses, place_of_death, year, month)
+
+#### Removing 2021-2022 rows ####
+cleaned_death_registry_data <- cleaned_death_registry_data[-c(746:841), ] 
+
+#### Recoding Months to be more Meaningful ####
+cleaned_death_registry_data <- cleaned_death_registry_data |> 
+  mutate( 
+    month = 
+      recode( 
+        month, 
+        "01" = "Jan", 
+        "02" = "Feb", 
+        "03" = "Mar",
+        "04" = "Apr",
+        "05" = "May",
+        "06" = "Jun",
+        "07" = "Jul",
+        "08" = "Aug",
+        "09" = "Sep",
+        "10" = "Oct",
+        "11" = "Nov",
+        "12" = "Dec"
+      ) 
+  )
+
+#### Recoding Civic Centres to be more Meaningful ####
+cleaned_death_registry_data <- cleaned_death_registry_data |> 
+  mutate( 
+    civic_centre = 
+      recode( 
+        civic_centre, 
+        "ET" = "Etobicoke", 
+        "NY" = "North York", 
+        "SC" = "Scarborough", 
+      ) 
+  )
+
[…]

### 2023-01-31 02-data_cleaning.R

@@ -1,49 +1,0 @@
-#### Preamble ####
-# Purpose: Clean the place of death data downloaded from Open Data Toronto's statistics on death registration. Link: https://open.toronto.ca/dataset/death-registry-statistics/
-# Author: Chloe Thierstein
-# Email: chloe.thierstein@utoronto.ca
-# Date: 23 January 2023
-# Prerequisites: Need to know where to find the City of Toronto death registry data
-
-#### workspace set-up ####
-#### install.packages(tidyverse) #####
-#### install.packages(janitor) #####
-
-library(tidyverse)
-library(janitor)
-
-#### Clean Names ####
-cleaned_place_of_death_data_2012 <-
-  clean_names(raw_death_registry_data)
-
-head(cleaned_place_of_death_data_2012)
-
-#### Simplifying Date #### 
-cleaned_place_of_death_data_2012 <-
-  cleaned_place_of_death_data_2012 |>
-  separate(
-    col = time_period,
-    into = c("Year", "Month"),
-    sep = "-"
-  ) |>
-  select(-Month) #remove month column as unnecessary
-
-#### Selecting Relevant Rows ####
-cleaned_place_of_death_data_2012 <- cleaned_place_of_death_data_2012[c(82:155),] #only keeping rows for 2012
-
-
-#### Selecting Relevant Columns####
-cleaned_place_of_death_data_2012 <-
-  cleaned_place_of_death_data_2012 |>
-  select(
-    death_licenses,
-    place_of_death
-  )
-
-
-
-
-
-
-
-

### 2023-01-31 scripts/00-download_data.R

@@ -1,0 +1,28 @@
+#### Preamble ####
+# Purpose: Read in data from Open Data Toronto's statistics on death registration. Link: https://open.toronto.ca/dataset/death-registry-statistics/
+# Author: Chloe Thierstein
+# Email: chloe.thierstein@utoronto.ca
+# Date: 23 January 2023
+# Prerequisites: Need to know where to find the City of Toronto death registry data
+
+#### workspace set-up ####
+#### install.packages(tidyverse) #####
+#### install.packages(janitor) #####
+
+library(tidyverse)
+library(janitor)
+
+#### Read in Data ####
+raw_death_registry_data <-
+  read_csv(
+    file = 
+      "https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/cba07a90-984b-42d2-9131-701c8c2a9788/resource/cc418325-b3d0-4afd-ae64-ccb9b71c549c/download/Death%20Registry%20Statistics%20Data.csv",
+    show_col_types = FALSE,
+    skip = 0
+  )
+
+write_csv(
+  x = raw_death_registry_data,
+  file = "death_registry.csv"
+)
+

### 2023-01-31 scripts/01-data_cleaning.R

@@ -1,86 +1,0 @@
-#### Preamble ####
-# Purpose: Read in data from Open Data Toronto and Clean the civic centre data downloaded from Open Data Toronto's statistics on death registration. Link: https://open.toronto.ca/dataset/death-registry-statistics/
-# Author: Chloe Thierstein
-# Email: chloe.thierstein@utoronto.ca
-# Date: 23 January 2023
-# Prerequisites: Need to know where to find the City of Toronto death registry data
-
-#### workspace set-up ####
-#### install.packages(tidyverse) #####
-#### install.packages(janitor) #####
-
-library(tidyverse)
-library(janitor)
-
-#### Read in Data ####
-raw_death_registry_data <-
-  read_csv(
-    file = 
-      "https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/cba07a90-984b-42d2-9131-701c8c2a9788/resource/cc418325-b3d0-4afd-ae64-ccb9b71c549c/download/Death%20Registry%20Statistics%20Data.csv",
-    show_col_types = FALSE,
-    skip = 0
-  )
-
-write_csv(
-  x = raw_death_registry_data,
-  file = "death_registry.csv"
-)
-
-
-#### Basic Cleaning ####
-raw_death_registry_data <-
-  read_csv(
-    file = "death_registry.csv",
-    show_col_types = FALSE
-  )
-
-
-#### Clean Names ####
-cleaned_civic_centre_data_2012 <-
-  clean_names(raw_death_registry_data)
-
-head(cleaned_civic_centre_data_2012)
-
-#### Simplifying Date #### 
-cleaned_civic_centre_data_2012 <-
-  cleaned_civic_centre_data_2012 |>
-  separate(
-    col = time_period,
-    into = c("Year", "Month"),
-    sep = "-"
-  ) |>
-  select(-Month) #remove month column as unnecessary
-
-
-#### Selecting Relevant Rows ####
-cleaned_civic_centre_data_2012 <- cleaned_civic_centre_data_2012[c(82:155),] #only keeping rows for 2012
-
-#### Recoding Civic Centres to be more Meaningful ####
-cleaned_civic_centre_data_2012 <- cleaned_civic_centre_data_2012 |> 
-  mutate( 
-    civic_centre = 
-      recode( 
-        civic_centre, 
-        "ET" = "Etobicoke", 
-        "NY" = "North York", 
[…]

### 2023-02-02 outputs/paper/paper.qmd
cleaned death registry data, added references to bib, updated paper writing and code, updated read me file, re-organized files, created simulation

--- 2023-01-30 outputs/paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Data on Death Registrations Illustrates Issues with Ontario's Records Maintenance"
 author: "Chloe Thierstein"
@@
 date-format: "D MMMM YYYY"
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "First sentence.
+Second sentence.
+Third sentence.
+Fourth sentence.
+- The abstract should be three or four sentences.
+The abstract must tell the reader the top-level finding.
+What is the one thing that we learn about the world because of this paper?"
 format: pdf
@@
 ```{r}
-#| include: FALSE
+#| include: false
 
-knitr::opts_chunk$set(echo = TRUE)
+#### Workspace Set-up ####
+
+#### install.packages(tidyverse) ####
+#### install.packages(janitor) ####
+#### install.packages(ggplot2) ####
+#### install.packages("knitr") ####
+#### install.packages("RColorBrewer") ####
+#### install.packages("kableExtra") ####
+#### install.packages("bookdown") ####
+#### install.packages("rmarkdown") ####
+#### install.packages("patchwork") ####
 
 library(tidyverse)
-library(palmerpenguins)
+library(opendatatoronto)
+library(janitor)
+library(ggplot2)
+library(dplyr)
+library(knitr)
+library(RColorBrewer)
+library(kableExtra)
+library(bookdown)
+library(patchwork)
+
 ```
@@
 # Introduction
+
+two or three paragraphs of content. and there should be an additional final paragraph that sets out the remainder of the paper
 
@@
 
+\-\-\-\-\-\-\-\-\--
+
+During the period of 2011 and 2020, it is estimated that on average just under 100 thousand people died each year in Ontario, Canada [@citestatscan1].
+The province of Ontario represents the location of a large portion of deaths in Canada as it takes on an estimated 37% of all deaths in Canada from 2011-2020 [@citestatscan1].
+From 2011 to 2020, the population in Ontario has grown from over 13 million to just under 15 million [@citestatscan2].
+Significantly, the city of Toronto makes up approximately 44% of the population of Ontario [@citestatscan2].
+
+\-\-\-\-\-\-\-\-\-\--
+
+In this paper I will be analyzing death registry data from four civic centers in the greater Toronto area, Etobicoke, North York, and Scarborough.
+First, I will compare rates of death registration from these civic centers over the period of 2011-2020.
+Next, I will compare rates of death registration against place of death in order to discuss the rate of deaths inside and outside the Toronto city limits.
+Finally, I will discuss trends in death registrations from each civic center.
+
+\-\-\-\-\-\-\-\-\--
+
 # Data
 
-Our data is of penguins (@fig-bills).
+data section should thoroughly and precisely discuss the source of the data and the bias this brings (ethical, statistical, and otherwise).
+comprehensively describe and summarize the data using text, graphs, and tables must be made iwth knitr or gt.
+graphs must show the actual data, or as close to it as possible, not summary statistics.
+(Graphs and tables should be cross-references in the text e.g., 'Table 1 shows...')
+
+\-\-\-\-\-\-\-\-\-\-\--
+
+Analysis for this paper utilizes R the statistical programming language [@citeR], along with several packages.
+These packages are, tidyverse [@citetidyverse], janitor [@citejanitor], and dplyr [@citedplyr].
+The data I am analyzing come from Open Data Toronto and it is imported using the opendatatoronto package [@citeopendatatoronto].
+**All figures in this paper have been created using ggplot2 [@citeggplot2] and the tables have been created with knitr [@citeknitr] and kableExtra [@citekablextra], packages.
+The color styles in my graph were created by using the RColorBrewer [@citercolorbrewer] package and any graph combinations were made using the Patchwork [@citepatchwork] package.**
+
+\-\-\-\-\-\-\-\-\-\-\--
+
+The data I am analyzing comes from Registry Services staff located in three of the civic centers (Etobicoke, North York, Scarborough) and
+
+\-\-\-\-\-\-\-\-\-\-\--
+
+"This dataset includes information relating to the registration of deaths documented by Registry Services staff located in four of the civic centres (Scarborough, North York, Toronto and Etobicoke).
+Registration of death data is entered into the Registry Services Tracking System (RSTS), from which aggregate statistical information is generated to create the dataset.
+The dataset resides in an Oracle database in the City's environment.
+
+\-\-\--
+
+The data I am analyzing comes from death registrations which are entered into the Registry Services Tracking System (RSTS) by Registry Services staff located at three of the civic centres (Etobicoke, North York, Scarborough) .
+
+@citeopendatatoronto2
+
+\-\-\--
+
+The dataset is created in support of the Vital Statistics Act, which is a Provincial legislation.
+The dataset also supports the City's operational requirements and business functions." [@citeopendatatoronto]
+
+"\_id = unique row identifier for Open Data database
+
+civic_centre = civic centre code
+
+death licenses = number of deaths registered in the month
+
+place_of_death = place where death occurred
+
+time_period = month death registered" [@citeopendatatoronto]
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\--
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
+#| include: FALSE
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+#### Read in Data ####
+
+raw_death_registry_data <-
+  read_csv(
+    file = 
+      "https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/cba07a90-984b-42d2-9131-701c8c2a9788/resource/cc418325-b3d0-4afd-ae64-ccb9b71c549c/download/Death%20Registry%20Statistics%20Data.csv",
+    show_col_types = FALSE,
+    skip = 0
+  )
+
+#### Creating a CSV File for Raw Data ####
+write_csv(
+  x = raw_death_registry_data,
+  file = "death_registry.csv" #create raw data .csv file and store it in inputs folder
+)
+
 ```
 
-Talk more about it.
+```{r}
+#| include: false
 
-Also bills and their average (@fig-billssssss).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)
+#### Data Cleaning ####
+
+#### Clean Names ####
+cleaned_death_registry_data <-
+  clean_names(raw_death_registry_data)
+
+#### Simplifying Date #### 
+cleaned_death_registry_data <-
+  cleaned_death_registry_data |>
+  separate(
+    col = time_period,
+    into = c("Year", "Month"),
+    sep = "-"
+  ) |>
+  select(civic_centre, death_licenses, place_of_death, Year, Month)
+
+#### Removing 2011-2015 and 2021-2023 Data for Relevance ####
+cleaned_death_registry_data <- cleaned_death_registry_data[-c(1:371,746:847), ]
+
+#### Recoding Months to be More Meaningful ####
+cleaned_death_registry_data <- cleaned_death_registry_data |> 
+  mutate( 
+    Month = 
+      recode( 
+        Month, #abbreviated months to better fit graphs withotu losing meaning
+        "01" = "Jan", 
+        "02" = "Feb", 
+        "03" = "Mar",
+        "04" = "Apr",
+        "05" = "May",
+        "06" = "Jun",
+        "07" = "Jul",
+        "08" = "Aug",
+        "09" = "Sep",
+        "10" = "Oct",
+        "11" = "Nov",
+        "12" = "Dec"
+      ) 
+  )
+
+#### Recoding Civic Centres to be More Meaningful ####
+cleaned_death_registry_data <- cleaned_death_registry_data |> 
+  mutate( 
+    civic_centre = 
+      recode( 
+        civic_centre, 
+        "ET" = "Etobicoke", 
+        "NY" = "North York", 
+        "SC" = "Scarborough", 
+      ) 
+  )
+
+#### Ordering Months #####
+cleaned_death_registry_data$Month <- factor(cleaned_death_registry_data$Month, levels=c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+#### Filtering Out "Toronto" Civic Centre from Data Due to Missing Data ####
+cleaned_death_registry_data <- cleaned_death_registry_data |>
+  filter(!grepl('TO', civic_centre))
+
+#### Creating a CSV File for Cleaned Data ####
+write_csv(
+  x = cleaned_death_registry_data,
+  file = "cleaned_death_registry.csv"  #create cleaned data .csv file and store it in inputs folder
+)
+
+```
+
+([@fig-deathlicensesbymonthandyear])
 
 ```{r}
-#| label: fig-billssssss
-#| fig-cap: More bills of penguins
 #| echo: false
-#| fig-width: 8
-#| fig-height: 4
+#| warning: false
+#| label: fig-deathlicensesbymonthandyear
+#| fig-cap: "Death Licenses Given Out by Month from 2016-2020 at Three Toronto Civic Centres"
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = FALSE) +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+#### Creating Graph to Compare Months Against Death Licenses Given Out From 2016-2020, Including Visualization of Years ####
+
+ggplot(data=cleaned_death_registry_data, aes(x = Month, y = death_licenses, fill = Year)) +
+  theme_classic() +
+  geom_bar(stat = "identity") +
+  labs(
+    x = "Month",
+    y = "Death Licenses"
+  ) +
+  scale_fill_brewer(palette = "Set3")
+
 ```
 
-Talk way more about it.
+```{r}
+#| echo: false
+#| warning: false
+#| label: fig-deathlicensesbyyearsolid
+#| fig-cap: "Death Licenses Given Out by All Civic Centres Yearly From 2016-2020 "
 
-# Model
+#### Creating Graph to Compare Years Against Death Licenses Given Out From 2016-2020 ####
 
-$$
-Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
-$$ {#eq-bayes}
+ggplot(data=cleaned_death_registry_data, aes(x = Year, y = death_licenses, fill = Year)) +
+  geom_bar(stat = "identity") +
+  theme_classic() + 
+  labs(
+    x = "Year",
+    y = "Death Licenses"
+  ) +
+  theme(legend.position = "none")+
+  scale_fill_brewer(palette="Set3")
+```
 
-@eq-bayes seems useful, eh?
+```{r}
+#| include: false
 
-Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR].
-We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.
+#### Creating Graph to Compare Months Against Death Licenses Given Out From 2016-2020 From Each Civic Centre ####
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+month_by_death_license <- ggplot(data=cleaned_death_registry_data, aes(x = Month, y = death_licenses, fill = civic_centre)) + #assign ggplot to new variable in order to use patchwork package later
+  geom_bar(stat = "identity") +
+  theme_classic() + 
+  labs(
+    x = "Month",
+    y = "Death Licenses",
+    fill = "Civic Centre"
+  ) +
+  scale_fill_brewer(palette="Set2")+
+  theme(legend.position = "none") + #remove legend as next graph will have it for better readability
+    theme(axis.title.y = element_text(size = 10))+
+    theme(axis.title.x = element_text(size = 10))
 
-# Results
+#### Creating Graph to Compare Years Against Death Licenses Given Out From 2016-2020 From Each Civic Centre ####
 
-# Discussion
+year_by_death_license <-  ggplot(data=cleaned_death_registry_data, aes(x = Year, y = death_licenses, fill = civic_centre)) + #assign ggplot to new variable in order to use patchwork package later
+  geom_bar(stat = "identity") +
+  theme_classic() + 
+  labs(
+    x = "Year",
+    y = "Death Licenses",
+    fill = "Civic Centre"
+  )+
+  scale_fill_brewer(palette="Set2")+
+  theme(legend.position = "bottom") + #include legend at bottom to ensure readability when using patchwork package
+  theme(axis.title.y = element_text(size = 10))+
+  theme(axis.title.x = element_text(size = 10))
+```
 
-## First discussion point
+```{r}
+#| echo: false
+#| warning: false
+#| label: fig-deathlicensesmonthly
+#| fig-cap: "Death Licenses Given Out by Each Civic Centre Yearly and Monthly Over the Period of 2016-2020"
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+#### Using Patchwork Package to Combine Two Graphs ####
 
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
+(month_by_death_license) / (year_by_death_license) +
+  plot_annotation(tag_levels = 'A') +
+  plot_layout(ncol = 1)
+```
 
@@
 
-\appendix
+```{r}
+#| include: false
 
-# Appendix {-}
+#### Creating Datasets for 5 Year Period by Month at Etobicoke Civic Centre ####
 
-# Additional details
+#### Repeat Process for Each Year (2016-2020) ####
+
+#### 2016 ####
+et_avg_num_deaths_by_month_2016 <-cleaned_death_registry_data |>
+  filter(Year == "2016" | civic_centre == "Etobicoke") |> #filter for year and civic centre
+  group_by(Month) |> #group by month
+  summarize("Average # of Deaths Registered" = mean(death_licenses)) #find average of death licenses given out
+
+#### 2017 ####
+et_avg_num_deaths_by_month_2017 <-cleaned_death_registry_data |>
+  filter(Year == "2017" | civic_centre == "Etobicoke") |>
+  group_by(Month) |> 
+  summarize("Average # of Deaths Registered" = mean(death_licenses))
+
+#### 2018 ####
+et_avg_num_deaths_by_month_2018 <-cleaned_death_registry_data |>
+  filter(Year == "2018" | civic_centre == "Etobicoke") |>
+  group_by(Month) |> 
+  summarize("Average # of Deaths Registered" = mean(death_licenses))
+
+#### 2019 ####
+et_avg_num_deaths_by_month_2019 <-cleaned_death_registry_data |>
+  filter(Year == "2019" | civic_centre == "Etobicoke") |>
+  group_by(Month) |> 
+  summarize("Average # of Deaths Registered" = mean(death_licenses))
+
+#### 2020 ####
+et_avg_num_deaths_by_month_2020 <-cleaned_death_registry_data|>
+  filter(Year == "2020" | civic_centre == "Etobicoke") |>
+  group_by(Month) |> 
+  summarize("Average # of Deaths Registered" = mean(death_licenses))
+
+```
+
+```{r}
+#| include: false
+
+#### Combining Datasets for 5 Year Period by Month at Etobicoke Civic Centre ####
+
+#### After 1st Instance (2016), Repeat 2nd Instance Process for Each Year (2017-2020) ####
+
+#### 2016 ####
+et_all_avg_num_deaths_by_month_data <- #new dataset
+  et_avg_num_deaths_by_month_2016 |> #start with first dataset in this spot
+  left_join(et_avg_num_deaths_by_month_2016, by = "Month") #use left_join() to combine dataset
+
+#### 2017 ####
+et_all_avg_num_deaths_by_month_data <- 
+  et_all_avg_num_deaths_by_month_data |> #for second and all following datasets use "new" dataset in this position
+  left_join(et_avg_num_deaths_by_month_2017, by = "Month")
+
+#### 2018 ####
+et_all_avg_num_deaths_by_month_data <- 
+  et_all_avg_num_deaths_by_month_data |>
+  left_join(et_avg_num_deaths_by_month_2018, by = "Month")
+
+#### 2019 ####
+et_all_avg_num_deaths_by_month_data <- 
+  et_all_avg_num_deaths_by_month_data |>
+  left_join(et_avg_num_deaths_by_month_2019, by = "Month")
+
+#### 2020 ####
+et_all_avg_num_deaths_by_month_data <-
+  et_all_avg_num_deaths_by_month_data |>
+  left_join(et_avg_num_deaths_by_month_2020, by = "Month")
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| label: tbl-avgnumofdeathsmonthet
+#| tbl-cap: "Average Number of Deaths Per Month from 2016-2020 at Etobicoke Centre"
+
+#### Creating a Table to Demonstrate Average Number of Deaths Per Month from 2016-2020 at Etobicoke Centre ####
+
+et_all_avg_num_deaths_by_month_data|>
+  select(Month, "Average # of Deaths Registered.y","Average # of Deaths Registered.x.x", "Average # of Deaths Registered.y.y", "Average # of Deaths Registered.x.x.x", "Average # of Deaths Registered.y.y.y")|> #previous step created a duplicate column, use select() to remove the second column in dataset
+     kable(
+    col.names = c( #rename columns to be more meaningful
+      "Month",
+      "2016",
+      "2017",
+      "2018",
+      "2019",
+      "2020"
+    ),
+    digits = 0,
+    booktabs = TRUE,
+    linesep = "") |>
+    add_header_above(c(" " = 1, "Average Number of Deaths" = 5)) |>
+  kable_styling(full_width = TRUE)
+
+```
+
+```{r}
+#| include: false
+
+#### Creating Datasets for 5 Year Period by Month at Scarborough Civic Centre ####
+
+#### After 1st Instance (2016), Repeat 2nd Instance Process for Each Year (2017-2020) ####
+
+#### 2016 ####
+sc_avg_num_deaths_by_month_2016 <-cleaned_death_registry_data |>
+  filter(Year == "2016" | civic_centre == "Scarborough") |> #filter for year and civic centre
+  group_by(Month) |>  #group by month
+  summarize("Average # of Deaths Registered" = mean(death_licenses)) #find average of death licenses given out
+
+#### 2017 ####
+sc_avg_num_deaths_by_month_2017 <-cleaned_death_registry_data |>
+  filter(Year == "2017" | civic_centre == "Scarborough") |>
+  group_by(Month) |> 
+  summarize("Average # of Deaths Registered" = mean(death_licenses))
+
+#### 2018 ####
+sc_avg_num_deaths_by_month_2018 <-cleaned_death_registry_data |>
+  filter(Year == "2018" | civic_centre == "Scarborough") |>
+  group_by(Month) |> 
+  summarize("Average # of Deaths Registered" = mean(death_licenses))
+
+#### 2019 ####
+sc_avg_num_deaths_by_month_2019 <-cleaned_death_registry_data |>
+  filter(Year == "2019" | civic_centre == "Scarborough") |>
+  group_by(Month) |> 
+  summarize("Average # of Deaths Registered" = mean(death_licenses))
+
+#### 2020 ####
+sc_avg_num_deaths_by_month_2020 <-cleaned_death_registry_data|>
+  filter(Year == "2020" | civic_centre == "Scarborough") |>
+  group_by(Month) |> 
+  summarize("Average # of Deaths Registered" = mean(death_licenses))
+
+```
+
+```{r}
+#| include: false
+
+#### Combining Datasets for 5 Year Period by Month at Scarborough Civic Centre ####
+
+#### After 1st Instance (2016), Repeat 2nd Instance Process for Each Year (2017-2020) ####
+
+#### 2016 ####
+sc_all_avg_num_deaths_by_month_data <- #new dataset
+  sc_avg_num_deaths_by_month_2016 |> #start with first dataset in this spot
+  left_join(sc_avg_num_deaths_by_month_2016, by = "Month") #use left_join() to combine dataset
+
+#### 2017 ####
+sc_all_avg_num_deaths_by_month_data <- 
+  sc_all_avg_num_deaths_by_month_data |> #for second and all following datasets use "new" dataset in this position
+  left_join(sc_avg_num_deaths_by_month_2017, by = "Month")
+
+#### 2018 ####
+sc_all_avg_num_deaths_by_month_data <- 
+  sc_all_avg_num_deaths_by_month_data |>
+  left_join(sc_avg_num_deaths_by_month_2018, by = "Month")
+
+#### 2019 ####
+sc_all_avg_num_deaths_by_month_data <- 
+  sc_all_avg_num_deaths_by_month_data |>
+  left_join(sc_avg_num_deaths_by_month_2019, by = "Month")
+
+#### 2020 ####
+sc_all_avg_num_deaths_by_month_data <-
+  sc_all_avg_num_deaths_by_month_data |>
+  left_join(sc_avg_num_deaths_by_month_2020, by = "Month")
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| label: tbl-avgnumofdeathsmonthsc
+#| tbl-cap: "Average Number of Deaths Per Month from 2016-2020 at Scarborough"
+
+#### Creating a Table to Demonstrate Average Number of Deaths Per Month From 2016-2020 at Scarborough Centre ####
+
+sc_all_avg_num_deaths_by_month_data|>
+  select(Month, "Average # of Deaths Registered.y","Average # of Deaths Registered.x.x", "Average # of Deaths Registered.y.y", "Average # of Deaths Registered.x.x.x", "Average # of Deaths Registered.y.y.y")|> #previous step created a duplicate column, use select() to remove the second column in dataset
+  kable(
+    col.names = c( #rename columns to be more meaningful
+      "Month",
+      "2016",
+      "2017",
+      "2018",
+      "2019",
+      "2020"
+    ),
+    digits = 0,
+    booktabs = TRUE,
+    linesep = "") |>
+  add_header_above(c(" " = 1, "Average Number of Deaths" = 5)) |>
+  kable_styling(full_width = TRUE)
+
+```
+
+```{r}
+#| include: false
+
+#### Creating Datasets for 5 Year Period by Month at North York Civic Centre ####
+
+#### Repeat Process for Each Year (2016-2020) ####
+
+#### 2016 ####
+ny_avg_num_deaths_by_month_2016 <-cleaned_death_registry_data |>
+  filter(Year == "2016" | civic_centre == "North York") |> #filter for year and civic centre
+  group_by(Month) |> #group by month
+  summarize("Average # of Deaths Registered" = mean(death_licenses)) #find average of death licenses given out
+
+#### 2017 ####
+ny_avg_num_deaths_by_month_2017 <-cleaned_death_registry_data |>
+  filter(Year == "2017" | civic_centre == "North York") |>
+  group_by(Month) |> 
+  summarize("Average # of Deaths Registered" = mean(death_licenses))
+
+#### 2018 ####
+ny_avg_num_deaths_by_month_2018 <-cleaned_death_registry_data |>
+  filter(Year == "2018" | civic_centre == "North York") |>
+  group_by(Month) |> 
+  summarize("Average # of Deaths Registered" = mean(death_licenses))
+
+#### 2019 ####
+ny_avg_num_deaths_by_month_2019 <-cleaned_death_registry_data |>
+  filter(Year == "2019" | civic_centre == "North York") |>
+  group_by(Month) |> 
+  summarize("Average # of Deaths Registered" = mean(death_licenses))
+
+#### 2020 ####
+ny_avg_num_deaths_by_month_2020 <-cleaned_death_registry_data|>
+  filter(Year == "2020" | civic_centre == "North York") |>
+  group_by(Month) |> 
+  summarize("Average # of Deaths Registered" = mean(death_licenses))
+
+```
+
+```{r}
+#| include: false
+
+#### Combining Datasets for 5 Year Period by Month at North York Civic Centre ####
+
+#### After 1st Instance (2016), Repeat 2nd Instance Process for Each Year (2017-2020) ####
+
+#### 2016 ####
+ny_all_avg_num_deaths_by_month_data <- #new dataset
+  ny_avg_num_deaths_by_month_2016 |> #start with first dataset in this spot
+  left_join(ny_avg_num_deaths_by_month_2016, by = "Month") #use left_join() to combine dataset
+
+#### 2017 ####
+ny_all_avg_num_deaths_by_month_data <- 
+  ny_all_avg_num_deaths_by_month_data |>#for second and all following datasets use "new" dataset in this position
+  left_join(ny_avg_num_deaths_by_month_2017, by = "Month")
+
+#### 2018 ####
+ny_all_avg_num_deaths_by_month_data <- 
+  ny_all_avg_num_deaths_by_month_data |>
+  left_join(ny_avg_num_deaths_by_month_2018, by = "Month")
+
+#### 2019 ####
+ny_all_avg_num_deaths_by_month_data <- 
+  ny_all_avg_num_deaths_by_month_data |>
+  left_join(ny_avg_num_deaths_by_month_2019, by = "Month")
+
+#### 2020 ####
+ny_all_avg_num_deaths_by_month_data <-
+  ny_all_avg_num_deaths_by_month_data |>
+  left_join(ny_avg_num_deaths_by_month_2020, by = "Month") 
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| label: tbl-avgnumofdeathsmonthny
+#| tbl-cap: "Average Number of Deaths Per Month from 2016-2020 at North York Centre"
+
+#### Creating a Table to Demonstrate Average Number of Deaths Per Month From 2016-2020 at North York Centre ####
+
+ny_all_avg_num_deaths_by_month_data|>
+  select(Month, "Average # of Deaths Registered.y","Average # of Deaths Registered.x.x", "Average # of Deaths Registered.y.y", "Average # of Deaths Registered.x.x.x", "Average # of Deaths Registered.y.y.y")|> #previous step created a duplicate column, use select() to remove the second column in dataset
+  kable(
+    col.names = c( #rename columns to be more meaningful
+      "Month",
+      "2016",
+      "2017",
+      "2018",
+      "2019",
+      "2020"
+    ),
+    digits = 0,
+    booktabs = TRUE,
+    linesep = "") |>
+  add_header_above(c(" " = 1, "Average Number of Deaths" = 5)) |>
+  kable_styling(full_width = TRUE)
+```
 

### 2023-02-02 01-data_cleaning.R

@@ -1,31 +1,50 @@
 
 #### Preamble ####
-# Purpose: Read in data from Open Data Toronto and clean the civic centre data from 2012 downloaded from Open Data Toronto's statistics on death registration. Link: https://open.toronto.ca/dataset/death-registry-statistics/
+# Purpose: Read in data from Open Data Toronto and clean the civic centre data from 2012 downloaded from Open Data Toronto's statistics on death registration.Link: https://open.toronto.ca/dataset/death-registry-statistics/
 # Author: Chloe Thierstein
 # Email: chloe.thierstein@utoronto.ca
 # Date: 23 January 2023
 # Prerequisites: Need to know where to find the City of Toronto death registry data
 
-#### workspace set-up ####
+#| include: false
+
+#### Workspace Set-up ####
+
 #### install.packages(tidyverse) ####
 #### install.packages(janitor) ####
+#### install.packages(ggplot2) ####
 #### install.packages("knitr") ####
+#### install.packages("RColorBrewer") ####
+#### install.packages("kableExtra") ####
 
-install.packages("RColorBrewer")
-library("RColorBrewer")
-
-library(knitr)
 library(tidyverse)
+library(opendatatoronto)
 library(janitor)
+library(ggplot2)
+library(dplyr)
+library(knitr)
+library(RColorBrewer)
+library(kableExtra)
 
 
+#### Read in Data ####
+raw_death_registry_data <-
+  read_csv(
+    file = 
+      "https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/cba07a90-984b-42d2-9131-701c8c2a9788/resource/cc418325-b3d0-4afd-ae64-ccb9b71c549c/download/Death%20Registry%20Statistics%20Data.csv",
+    show_col_types = FALSE,
+    skip = 0
+  )
+
+write_csv(
+  x = raw_death_registry_data,
+  file = "death_registry.csv"
+)
 
 
 #### Clean Names ####
 cleaned_death_registry_data <-
   clean_names(raw_death_registry_data)
-
-head(cleaned_death_registry_data)
 
 
 #### Simplifying Date #### 
@@ -33,20 +52,21 @@ cleaned_death_registry_data <-
   cleaned_death_registry_data |>
   separate(
     col = time_period,
-    into = c("year", "month"),
+    into = c("Year", "Month"),
     sep = "-"
[…]

### 2023-02-02 scripts/00-simulation.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Simulate data from Open Data Toronto's statistics on death registration. Link: https://open.toronto.ca/dataset/death-registry-statistics/
+# Purpose: Simulate and test data from Open Data Toronto's statistics on death registration. Link: https://open.toronto.ca/dataset/death-registry-statistics/
 # include columns ID, Death Licenses, Civic Centre, Place of Death, Month (as placeholder for date)
 # Author: Chloe Thierstein
 # Email: chloe.thierstein@mail.utoronto.ca
@@ -13,18 +13,21 @@ library(tibble)
 
 simulated_data <-
   tibble(
+    #### column 1: Unique ID: generated by random number between 841 
     "ID" = 1:841,
+    #### column 2: Death Licenses: generated by random number between 1 and 1000
     "Death Licenses" = runif(n = 841, min = 1, max = 1000),
+    #### column 3: Civic Centres: samples 3 centers with size of 841 and it can have duplicates
     "Civic Centre" = sample(
       x = c(
         "Etobicoke",
         "North York",
-        "Scarborough",
-        "Toronto"
+        "Scarborough"
      ),
      size = 841,
      replace = TRUE
     ),
+    #### column 4: place of death: samples 2 options for place of death with size of 841 and it can have duplicates
     "Place of Death" = sample(
       x = c(
         "Toronto",
@@ -32,7 +35,19 @@ simulated_data <-
       ),
       size = 841,
       replace = TRUE
+    ),#### column 5: Year: Samples 5 options for year with size of 841 and it can have duplicates
+    "Year" = sample(
+      x = c(
+        "2016",
+        "2017",
+        "2018",
+        "2019",
+        "2020"
+      ),
+      size = 841,
+      replace = TRUE
     ),
+    #### column 6: month: Samples 12 options for month with size of 841 and it can have duplicates
     "Month" = sample(
       x = c(
         "January",
@@ -53,3 +68,19 @@ simulated_data <-
   )
 )
 simulated_data
+
+#### testing simulated data ####
+
+#### test 1 #### 
+
+simulated_data$"Civic Centre" |>
+  unique() |>
+  length() == 3
+
+#### test 2 #### 
+
+simulated_data$"Year" |> min() == 2016
+
+#### test 3 ####
+
+simulated_data$"Death Licenses" |> class() == "numeric"

### 2023-02-02 outputs/paper/paper.qmd
updated re-organized scripts, fixed import of data to quarto file,

@@
 
-#### Workspace Set-up ####
-
-#### install.packages(tidyverse) ####
-#### install.packages(janitor) ####
-#### install.packages(ggplot2) ####
-#### install.packages("knitr") ####
-#### install.packages("RColorBrewer") ####
-#### install.packages("kableExtra") ####
-#### install.packages("bookdown") ####
-#### install.packages("rmarkdown") ####
-#### install.packages("patchwork") ####
+#### workspace set-up ####
 
@@
 library(patchwork)
+library(readr)
+library(here)
 
+#### import cleaned_death_registry.csv ####
+cleaned_death_registry_data <- read.csv(here::here("/Users/chloethierstein/Desktop/death_registry_analysis/inputs/data/cleaned_death_registry.csv"))
+
+#### turning years into factor ####
+cleaned_death_registry_data$Year <- as.factor(cleaned_death_registry_data$Year)
 ```
@@
 The data I am analyzing come from Open Data Toronto and it is imported using the opendatatoronto package [@citeopendatatoronto].
-**All figures in this paper have been created using ggplot2 [@citeggplot2] and the tables have been created with knitr [@citeknitr] and kableExtra [@citekablextra], packages.
+**
+
+All figures in this paper have been created using ggplot2 [@citeggplot2] and the tables have been created with knitr [@citeknitr] and kableExtra [@citekablextra], packages.
 The color styles in my graph were created by using the RColorBrewer [@citercolorbrewer] package and any graph combinations were made using the Patchwork [@citepatchwork] package.**
@@
 
-```{r}
-#| include: FALSE
-
-#### Read in Data ####
-
-raw_death_registry_data <-
-  read_csv(
-    file = 
-      "https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/cba07a90-984b-42d2-9131-701c8c2a9788/resource/cc418325-b3d0-4afd-ae64-ccb9b71c549c/download/Death%20Registry%20Statistics%20Data.csv",
-    show_col_types = FALSE,
-    skip = 0
-  )
-
-#### Creating a CSV File for Raw Data ####
-write_csv(
-  x = raw_death_registry_data,
-  file = "death_registry.csv" #create raw data .csv file and store it in inputs folder
-)
-
-```
-
-```{r}
-#| include: false
-
-#### Data Cleaning ####
-
-#### Clean Names ####
-cleaned_death_registry_data <-
-  clean_names(raw_death_registry_data)
-
-#### Simplifying Date #### 
-cleaned_death_registry_data <-
-  cleaned_death_registry_data |>
-  separate(
-    col = time_period,
-    into = c("Year", "Month"),
-    sep = "-"
-  ) |>
-  select(civic_centre, death_licenses, place_of_death, Year, Month)
-
-#### Removing 2011-2015 and 2021-2023 Data for Relevance ####
-cleaned_death_registry_data <- cleaned_death_registry_data[-c(1:371,746:847), ]
-
-#### Recoding Months to be More Meaningful ####
-cleaned_death_registry_data <- cleaned_death_registry_data |> 
-  mutate( 
-    Month = 
-      recode( 
-        Month, #abbreviated months to better fit graphs withotu losing meaning
-        "01" = "Jan", 
-        "02" = "Feb", 
-        "03" = "Mar",
-        "04" = "Apr",
-        "05" = "May",
-        "06" = "Jun",
-        "07" = "Jul",
-        "08" = "Aug",
-        "09" = "Sep",
-        "10" = "Oct",
-        "11" = "Nov",
-        "12" = "Dec"
-      ) 
-  )
-
-#### Recoding Civic Centres to be More Meaningful ####
-cleaned_death_registry_data <- cleaned_death_registry_data |> 
-  mutate( 
-    civic_centre = 
-      recode( 
-        civic_centre, 
-        "ET" = "Etobicoke", 
-        "NY" = "North York", 
-        "SC" = "Scarborough", 
-      ) 
-  )
-
-#### Ordering Months #####
-cleaned_death_registry_data$Month <- factor(cleaned_death_registry_data$Month, levels=c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
-
-#### Filtering Out "Toronto" Civic Centre from Data Due to Missing Data ####
-cleaned_death_registry_data <- cleaned_death_registry_data |>
-  filter(!grepl('TO', civic_centre))
-
-#### Creating a CSV File for Cleaned Data ####
-write_csv(
-  x = cleaned_death_registry_data,
-  file = "cleaned_death_registry.csv"  #create cleaned data .csv file and store it in inputs folder
-)
-
-```
-
-([@fig-deathlicensesbymonthandyear])
+(@fig-deathlicensesbymonthandyear)
 
@@
 #| include: false
+#| message: false
 
@@
   scale_fill_brewer(palette="Set2")+
-  theme(legend.position = "none") + #remove legend as next graph will have it for better readability
-    theme(axis.title.y = element_text(size = 10))+
+  theme(legend.position = "none") + #include legend at bottom to ensure readability when using patchwork package
+  theme(axis.title.y = element_text(size = 10)) +
   theme(axis.title.x = element_text(size = 10))
+
+```
+
+```{r}
+#| include: false
+#| message: false
 
@@
 ```{r}
+#| warning: false
 #| echo: false
-#| warning: false
 #| label: tbl-avgnumofdeathsmonthet
@@
 
-et_all_avg_num_deaths_by_month_data|>
-  select(Month, "Average # of Deaths Registered.y","Average # of Deaths Registered.x.x", "Average # of Deaths Registered.y.y", "Average # of Deaths Registered.x.x.x", "Average # of Deaths Registered.y.y.y")|> #previous step created a duplicate column, use select() to remove the second column in dataset
+et_all_avg_num_deaths_by_month_data |>
+  select(
+    Month,
+    "Average # of Deaths Registered.y",
+    "Average # of Deaths Registered.x.x",
+    "Average # of Deaths Registered.y.y",
+    "Average # of Deaths Registered.x.x.x",
+    "Average # of Deaths Registered.y.y.y"
+  ) |> #previous step created a duplicate column, use select() to remove the second column in dataset
   kable(
-    col.names = c( #rename columns to be more meaningful
+    col.names = c(#rename columns to be more meaningful
       "Month",
@@
       "2019",
-      "2020"
-    ),
+      "2020"),
     digits = 0,
     booktabs = TRUE,
-    linesep = "") |>
-    add_header_above(c(" " = 1, "Average Number of Deaths" = 5)) |>
-  kable_styling(full_width = TRUE)
+    linesep = ""
+  ) |>
+  add_header_above(c(" " = 1, "Average Number of Deaths" = 5))
 
@@
 #| label: tbl-avgnumofdeathsmonthsc
-#| tbl-cap: "Average Number of Deaths Per Month from 2016-2020 at Scarborough"
+#| tbl-cap: "Average Number of Deaths Per Month from 2016-2020 at Scarborough Centre"
 
@@
     linesep = "") |>
-  add_header_above(c(" " = 1, "Average Number of Deaths" = 5)) |>
-  kable_styling(full_width = TRUE)
+  add_header_above(c(" " = 1, "Average Number of Deaths" = 5))
 
@@
     linesep = "") |>
-  add_header_above(c(" " = 1, "Average Number of Deaths" = 5)) |>
-  kable_styling(full_width = TRUE)
+  add_header_above(c(" " = 1, "Average Number of Deaths" = 5)) 
 ```
 
-\newpage
+\clearpage
 

### 2023-02-02 scripts/00-simulation.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Simulate and test data from Open Data Toronto's statistics on death registration. Link: https://open.toronto.ca/dataset/death-registry-statistics/
+# Purpose: Simulate from Open Data Toronto's statistics on death registration. Link: https://open.toronto.ca/dataset/death-registry-statistics/
 # include columns ID, Death Licenses, Civic Centre, Place of Death, Month (as placeholder for date)
 # Author: Chloe Thierstein
 # Email: chloe.thierstein@mail.utoronto.ca

### 2023-02-02 scripts/01-simulation_tests.R

@@ -1,0 +1,23 @@
+#### Preamble ####
+# Purpose: Test data from Open Data Toronto's statistics on death registration. Link: https://open.toronto.ca/dataset/death-registry-statistics/
+# include columns ID, Death Licenses, Civic Centre, Place of Death, Month (as placeholder for date)
+# Author: Chloe Thierstein
+# Email: chloe.thierstein@mail.utoronto.ca
+# Date: 23 January 2023
+# Prerequisites: Need to know where to find the City of Toronto death registry data
+
+#### testing simulated data ####
+
+#### test 1 #### 
+
+simulated_data$"Civic Centre" |>
+  unique() |>
+  length() == 3
+
+#### test 2 #### 
+
+simulated_data$"Year" |> min() == 2016
+
+#### test 3 ####
+
+simulated_data$"Death Licenses" |> class() == "numeric"

### 2023-02-02 scripts/02-download_data.R

@@ -5,14 +5,8 @@
 # Date: 23 January 2023
 # Prerequisites: Need to know where to find the City of Toronto death registry data
 
-#### workspace set-up ####
-#### install.packages(tidyverse) #####
-#### install.packages(janitor) #####
-
-library(tidyverse)
-library(janitor)
-
 #### Read in Data ####
+
 raw_death_registry_data <-
   read_csv(
     file = 
@@ -21,8 +15,9 @@ raw_death_registry_data <-
     skip = 0
   )
 
+#### Creating a CSV File for Raw Data ####
 write_csv(
   x = raw_death_registry_data,
-  file = "death_registry.csv"
+  file = "death_registry.csv" #create raw data .csv file and store it in inputs folder
 )
 

### 2023-02-02 scripts/03-data_cleaning.R

@@ -6,8 +6,6 @@
 # Date: 23 January 2023
 # Prerequisites: Need to know where to find the City of Toronto death registry data
 
-#| include: false
-
 #### Workspace Set-up ####
 
 #### install.packages(tidyverse) ####
@@ -16,6 +14,10 @@
 #### install.packages("knitr") ####
 #### install.packages("RColorBrewer") ####
 #### install.packages("kableExtra") ####
+#### install.packages("bookdown") ####
+#### install.packages("rmarkdown") ####
+#### install.packages("patchwork") ####
+#### install.packages("here") ####
 
 library(tidyverse)
 library(opendatatoronto)
@@ -25,27 +27,16 @@ library(dplyr)
 library(knitr)
 library(RColorBrewer)
 library(kableExtra)
+library(bookdown)
+library(patchwork)
+library(readr)
+library(here)
 
-
-#### Read in Data ####
-raw_death_registry_data <-
-  read_csv(
-    file = 
-      "https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/cba07a90-984b-42d2-9131-701c8c2a9788/resource/cc418325-b3d0-4afd-ae64-ccb9b71c549c/download/Death%20Registry%20Statistics%20Data.csv",
-    show_col_types = FALSE,
-    skip = 0
-  )
-
-write_csv(
-  x = raw_death_registry_data,
-  file = "death_registry.csv"
-)
-
+#### Data Cleaning ####
 
 #### Clean Names ####
 cleaned_death_registry_data <-
   clean_names(raw_death_registry_data)
-
 
 #### Simplifying Date #### 
 cleaned_death_registry_data <-
@@ -57,16 +48,15 @@ cleaned_death_registry_data <-
   ) |>
   select(civic_centre, death_licenses, place_of_death, Year, Month)
 
-#### Removing 2021-2023 data ####
+#### Removing 2011-2015 and 2021-2023 Data for Relevance ####
 cleaned_death_registry_data <- cleaned_death_registry_data[-c(1:371,746:847), ] 
 
-
-#### Recoding Months to be more Meaningful ####
+#### Recoding Months to be More Meaningful ####
 cleaned_death_registry_data <- cleaned_death_registry_data |> 
   mutate( 
     Month = 
       recode( 
-        Month, 
+        Month, #abbreviated months to better fit graphs withotu losing meaning
         "01" = "Jan", 
         "02" = "Feb", 
         "03" = "Mar",
[…]

### 2023-02-03 outputs/paper/paper.qmd
continued to work on and develop write-up for paper, re-organized files, failed to stop floating in figures in .qmd file <- need to look into that further

@@
 ---
-title: "Data on Death Registrations Illustrates Issues with Ontario's Records Maintenance"
+title: "Death Registrations May Indicate a Discreptency Between Expected and Reported Data During the COVID-19 Pandemic"
+subtitle: "An Analysis of Death Registrations from 2016-2020"
 author: "Chloe Thierstein"
 thanks: "Code and data are available at: https://github.com/cthierst/death_registry_analysis.git."
-date: "24 January 2023"
+date: "`r Sys.time()`"
 date-format: "D MMMM YYYY"
-abstract: "First sentence.
-Second sentence.
-Third sentence.
-Fourth sentence.
-- The abstract should be three or four sentences.
-The abstract must tell the reader the top-level finding.
-What is the one thing that we learn about the world because of this paper?"
+abstract: "The COVID-19 pandemic marked a significant shift in Canadian mortality rates in the year 2020.
+Using death registry data from three civic centres in the greater Toronto are from 2016 to 2020, we demonstrate that even data from a considerably reliable source may become distorted"
 format: pdf
 bibliography: references.bib
+fig-cap-location: top
 ---
+
+# 1 Introduction
+
+Canadian death registry statistics offer a useful tool in helping better our healthcare system as by monitoring trends in public health such as infectious diseases, suicide, and unintentional injuries they help the healthcare sector provide better services and resources, like screening and prevention programs [@citestatscan3].
+However, maintaining this data may not always be done accurately which could cause confusion over our understanding of Canadian demography and health.
+
+In this paper we look at death registry statistics from three civic centres in the greater Toronto area, Etobicoke, Scarborough and North York between 2016 and 2020.
+From this data, we find that the death licenses provided by these centres during the first year of the COVID-19 pandemic (2020), are much lower than in previous years [@citeopendatatoronto2].
+This is significant as Canada experienced an increase of 7.7% in deaths in 2020 [@citestatscan4], and 5.2% more deaths than would typically be expected when taking into account Canada's aging population [@citestatscan5].
+We reason that this discrepancy in expected data versus reported data could be the result of strain and complications put onto record-keeping during this time similarly to what was seen in early 2019 with long delays in the distribution of death, birth and marriage certificates due to high online demand [@citectvnews].
+Future work could look specifically at how strains on data reporting centres during the COVID-19 pandemic influenced data.
+
+This paper will begin with an overview of its data management, source and cleaning.
+Next we will briefly overview the drop off in trend from previous years in death registrations to begin our discussion.
+Next we will consider the relationship between death licenses provided at each civic centre over the period of 2016 to 2020, to better understand how the year 2020 differed from past years.
+Next, we discuss in more detail, death licenses provided by each civic centre to determine a possible cause for the uncharacteristic decline of death licenses in 2020.
+We will then compare the mean number of deaths from each civic centre to further gauge how these centres relate to rates of death registrations.
+Finally, we will consider the limitations of the data, including biases.
+
+# 2 Data
+
+## 2.1 Data Management
+
+This paper utilizes the R statistical programming language [@citeR], along with several packages.
+These packages are, tidyverse [@citetidyverse], janitor [@citejanitor], here [@citehere] and dplyr [@citedplyr].
+The data being analyzed comes from Open Data Toronto and it is imported using the opendatatoronto package [@citeopendatatoronto].
+All figures have been created using ggplot2 [@citeggplot2] and the tables have been created with knitr [@citeknitr] and kableExtra [@citekablextra], packages.
+The color styles in the graphs were created by using the RColorBrewer [@citercolorbrewer] package and any graph combinations were made using the Patchwork [@citepatchwork] package.
+
+## 2.2 Data Source and Cleaning
+
+The data comes from death registrations which are entered into the Registry Services Tracking System (RSTS) by Registry Services staff who are located at three of the civic centres, Etobicoke, North York and Scarborough [@citeopendatatoronto2].
+It's creation supports the Vital Statistics Act, a Provincial legislation [@citeopendatatoronto2] which involve the collection of deaths, marriages, stillbirths, and live births [@citestatscan3].
+This data set is updated monthly [@citeopendatatoronto2].
+The variables from this data set represent the civic centres; "Etobicoke", "North York", and "Scarborough", number of death licenses registered in the month, place of death; "Outside City Limits" or "Toronto", and time period by month and year in which the death was registered [@citeopendatatoronto2].
+
+To properly analyze this data for the purposes of this paper certain data was removed including, all data for the years 2011-2015 and 2021-2023.
+This was done to ensure that the data set being worked with had sufficient and consistent data to draw from without being too broad in scope.
+Additionally, rows from the civic centre variable labelled "Toronto" were removed as death registrations at this location were especially low and its data input, sporadic.
+Finally, the variable for place of death was not included in the analysis of this data as it was not relevant to its narrative.
+
+## 2.3 Data Analysis
 
@@
 
-#### workspace set-up ####
+#### Workspace Set-Up ####
 
+#### Load Library ####
 library(tidyverse)
@@
 library(kableExtra)
-library(bookdown)
 library(patchwork)
@@
 
-#### import cleaned_death_registry.csv ####
+#### Import cleaned_death_registry.csv ####
 cleaned_death_registry_data <- read.csv(here::here("/Users/chloethierstein/Desktop/death_registry_analysis/inputs/data/cleaned_death_registry.csv"))
 
-#### turning years into factor ####
+#### Turning Years into Factors ####
 cleaned_death_registry_data$Year <- as.factor(cleaned_death_registry_data$Year)
+
 ```
 
-# Introduction
-
-two or three paragraphs of content. and there should be an additional final paragraph that sets out the remainder of the paper
-
-You can and should cross-reference sections and sub-sections.
-For instance, Section \@ref(data).
-Quarto automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).
-
-\-\-\-\-\-\-\-\-\--
-
-During the period of 2011 and 2020, it is estimated that on average just under 100 thousand people died each year in Ontario, Canada [@citestatscan1].
-The province of Ontario represents the location of a large portion of deaths in Canada as it takes on an estimated 37% of all deaths in Canada from 2011-2020 [@citestatscan1].
-From 2011 to 2020, the population in Ontario has grown from over 13 million to just under 15 million [@citestatscan2].
-Significantly, the city of Toronto makes up approximately 44% of the population of Ontario [@citestatscan2].
-
-\-\-\-\-\-\-\-\-\-\--
-
-In this paper I will be analyzing death registry data from four civic centers in the greater Toronto area, Etobicoke, North York, and Scarborough.
-First, I will compare rates of death registration from these civic centers over the period of 2011-2020.
-Next, I will compare rates of death registration against place of death in order to discuss the rate of deaths inside and outside the Toronto city limits.
-Finally, I will discuss trends in death registrations from each civic center.
-
-\-\-\-\-\-\-\-\-\--
-
-# Data
-
-data section should thoroughly and precisely discuss the source of the data and the bias this brings (ethical, statistical, and otherwise).
-comprehensively describe and summarize the data using text, graphs, and tables must be made iwth knitr or gt.
-graphs must show the actual data, or as close to it as possible, not summary statistics.
-(Graphs and tables should be cross-references in the text e.g., 'Table 1 shows...')
-
-\-\-\-\-\-\-\-\-\-\-\--
-
-Analysis for this paper utilizes R the statistical programming language [@citeR], along with several packages.
-These packages are, tidyverse [@citetidyverse], janitor [@citejanitor], and dplyr [@citedplyr].
-The data I am analyzing come from Open Data Toronto and it is imported using the opendatatoronto package [@citeopendatatoronto].
-**
-
-All figures in this paper have been created using ggplot2 [@citeggplot2] and the tables have been created with knitr [@citeknitr] and kableExtra [@citekablextra], packages.
-The color styles in my graph were created by using the RColorBrewer [@citercolorbrewer] package and any graph combinations were made using the Patchwork [@citepatchwork] package.**
-
-\-\-\-\-\-\-\-\-\-\-\--
-
-The data I am analyzing comes from Registry Services staff located in three of the civic centers (Etobicoke, North York, Scarborough) and
-
-\-\-\-\-\-\-\-\-\-\-\--
-
-"This dataset includes information relating to the registration of deaths documented by Registry Services staff located in four of the civic centres (Scarborough, North York, Toronto and Etobicoke).
-Registration of death data is entered into the Registry Services Tracking System (RSTS), from which aggregate statistical information is generated to create the dataset.
-The dataset resides in an Oracle database in the City's environment.
-
-\-\-\--
-
-The data I am analyzing comes from death registrations which are entered into the Registry Services Tracking System (RSTS) by Registry Services staff located at three of the civic centres (Etobicoke, North York, Scarborough) .
-
-@citeopendatatoronto2
-
-\-\-\--
-
-The dataset is created in support of the Vital Statistics Act, which is a Provincial legislation.
-The dataset also supports the City's operational requirements and business functions." [@citeopendatatoronto]
-
-"\_id = unique row identifier for Open Data database
-
-civic_centre = civic centre code
-
-death licenses = number of deaths registered in the month
-
-place_of_death = place where death occurred
-
-time_period = month death registered" [@citeopendatatoronto]
-
-\-\-\-\-\-\-\-\-\-\-\-\-\-\--
-
-(@fig-deathlicensesbymonthandyear)
+The province of Ontario represents a large portion of deaths in Canada, taking on an estimated 37% of Canadian mortality from 2016-2020 [@citestatscan1].
+It is also estimated that on average, that just over 100 thousand people have died each year in Ontario between 2016 and 2020 [@citestatscan1].
+We may suggest then, that Toronto makes up a large portion of this mortality rate, as it makes up approximately 44% of the Ontarian population [@citestatscan2].
+Additionally, Canada has an aging population, meaning that it's mortality rate is expected to rise every year although it rose exponentially due to the COVID-19 pandemic [@citestatscan4].
+Therefore, we may suspect that the COVID-19 pandemic would increase the mortality rate significantly in Toronto as Canada saw an increase of 7.7% in deaths in 2020 [@citestatscan4].
 
@@
 #| warning: false
-#| label: fig-deathlicensesbymonthandyear
-#| fig-cap: "Death Licenses Given Out by Month from 2016-2020 at Three Toronto Civic Centres"
-
-#### Creating Graph to Compare Months Against Death Licenses Given Out From 2016-2020, Including Visualization of Years ####
-
-ggplot(data=cleaned_death_registry_data, aes(x = Month, y = death_licenses, fill = Year)) +
-  theme_classic() +
-  geom_bar(stat = "identity") +
-  labs(
-    x = "Month",
-    y = "Death Licenses"
-  ) +
-  scale_fill_brewer(palette = "Set3")
-
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| label: fig-deathlicensesbyyearsolid
-#| fig-cap: "Death Licenses Given Out by All Civic Centres Yearly From 2016-2020 "
+#| label: fig-1
+#| fig-cap: "Death Licenses Provided by All Civic Centres Yearly From 2016-2020 "
 
@@
   scale_fill_brewer(palette="Set3")
+
 ```
+
+When looking at @fig-1, however, we can see that from 2016 to 2018 to there was an increase in death licenses registered, as expected, followed by a slight drop in 2019.
+This drop in 2019 may be indicative of a backlog from Ontario's Vital Statistics in early 2019 which was cited as having delays of over three months, waiting for death, birth and marriage certificates [@citectvnews].
+What is most important in @fig-1 is that when looking at the "2020" bar in @fig-1, we can see that the death licenses registered drops significantly in comparison to any of the previous 4 months.
+
+```{r}
+#| echo: false
+#| warning: false
+#| label: fig-2
+#| fig-cap: "Death Licenses Provided by Month from 2016-2020 at Three Toronto Civic Centres"
+
+#### Creating Graph to Compare Months Against Death Licenses Given Out From 2016-2020, Including Visualization of Years ####
+
+#### Ordering Months #####
+cleaned_death_registry_data$Month <- factor(cleaned_death_registry_data$Month, levels=c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+ggplot(data=cleaned_death_registry_data, aes(x = Month, y = death_licenses, fill = Year)) +
+  theme_classic() +
+  geom_bar(stat = "identity") +
+  labs(
+    x = "Month",
+    y = "Death Licenses"
+  ) +
+  scale_fill_brewer(palette = "Set3")
+```
+
+when looking at @fig-2 we can see - death licenses provided by month and year from all three civic centres - highlights the spike in death licenses provided in May - this may indicate a spike in deaths near the beginning of COVID-19 pandemic, - \~need to find source to make this make sense\~
 
@@
 #| warning: false
-#| label: fig-deathlicensesmonthly
-#| fig-cap: "Death Licenses Given Out by Each Civic Centre Yearly and Monthly Over the Period of 2016-2020"
+#| label: fig-3
+#| fig-cap: "Death Licenses Provided by Each Civic Centre Yearly and Monthly from 2016-2020"
 
@@
 
-\newpage
+Looking briefly at @fig-3 we can see a surprising drop in death licenses provided in 2020.
+\###### this is surprising as death rates are said to have increased during the pandemic, yet there is a signficant decline \###### why?
+explain...
 
@@
 #| echo: false
-#| label: tbl-avgnumofdeathsmonthet
+#| label: tbl-1
 #| tbl-cap: "Average Number of Deaths Per Month from 2016-2020 at Etobicoke Centre"
@@
 #| warning: false
-#| label: tbl-avgnumofdeathsmonthsc
+#| label: tbl-2
 #| tbl-cap: "Average Number of Deaths Per Month from 2016-2020 at Scarborough Centre"
@@
 #| warning: false
-#| label: tbl-avgnumofdeathsmonthny
+#| label: tbl-3
 #| tbl-cap: "Average Number of Deaths Per Month from 2016-2020 at North York Centre"

### 2023-02-03 outputs/paper/paper.tex
continued to work on and develop write-up for paper, re-organized files, failed to stop floating in figures in .qmd file <- need to look into that further

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading,figureheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Death Registrations May Indicate a Discreptency Between Expected and Reported Data During the COVID-19 Pandemic},
  pdfauthor={Chloe Thierstein},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Death Registrations May Indicate a Discreptency Between Expected
and Reported Data During the COVID-19 Pandemic\thanks{Code and data are
available at:
https://github.com/cthierst/death\_registry\_analysis.git.}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{An Analysis of Death Registrations from 2016-2020}
\author{Chloe Thierstein}
\date{3 February 2023}

\begin{document}
\maketitle
\begin{abstract}
The COVID-19 pandemic marked a significant shift in Canadian mortality
rates in the year 2020. Using death registry data from three civic
centres in the greater Toronto are from 2016 to 2020, we demonstrate
that even data from a considerably reliable source may become distorted
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, sharp corners]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{1 Introduction}\label{introduction}}

Canadian death registry statistics offer a useful tool in helping better
our healthcare system as by monitoring trends in public health such as
infectious diseases, suicide, and unintentional injuries they help the
healthcare sector provide better services and resources, like screening
and prevention programs (Statistics Canada 2022b). However, maintaining
this data may not always be done accurately which could cause confusion
over our understanding of Canadian demography and health.

In this paper we look at death registry statistics from three civic
centres in the greater Toronto area, Etobicoke, Scarborough and North
York between 2016 and 2020. From this data, we find that the death
licenses provided by these centres during the first year of the COVID-19
pandemic (2020), are much lower than in previous years (Open Data
Toronto 2023). This is significant as Canada experienced an increase of
7.7\% in deaths in 2020 (Statistics Canada 2022a), and 5.2\% more deaths
than would typically be expected when taking into account Canada's aging
population (Statistics Canada 2021). We reason that this discrepancy in
expected data versus reported data could be the result of strain and
complications put onto record-keeping during this time similarly to what
was seen in early 2019 with long delays in the distribution of death,
birth and marriage certificates due to high online demand (Jeffords
2019). Future work could look specifically at how strains on data
reporting centres during the COVID-19 pandemic influenced data.

This paper will begin with an overview of its data management, source
and cleaning. Next we will briefly overview the drop off in trend from
previous years in death registrations to begin our discussion. Next we
will consider the relationship between death licenses provided at each
civic centre over the period of 2016 to 2020, to better understand how
the year 2020 differed from past years. Next, we discuss in more detail,
death licenses provided by each civic centre to determine a possible
cause for the uncharacteristic decline of death licenses in 2020. We
will then compare the mean number of deaths from each civic centre to
further gauge how these centres relate to rates of death registrations.
Finally, we will consider the limitations of the data, including biases.

\hypertarget{data}{%
\section{2 Data}\label{data}}

\hypertarget{data-management}{%
\subsection{2.1 Data Management}\label{data-management}}

This paper utilizes the R statistical programming language (R Core Team
2020), along with several packages. These packages are, tidyverse
(Wickham et al. 2019), janitor (Firke 2021), here (Müller 2020) and
dplyr (Wickham et al. 2022). The data being analyzed comes from Open
Data Toronto and it is imported using the opendatatoronto package
(Gelfand 2022). All figures have been created using ggplot2 (Wickham
2016) and the tables have been created with knitr (Xie 2023) and
kableExtra (Zhu 2021), packages. The color styles in the graphs were
created by using the RColorBrewer (Neuwirth 2022) package and any graph
combinations were made using the Patchwork (Pedersen 2022) package.

\hypertarget{data-source-and-cleaning}{%
\subsection{2.2 Data Source and
Cleaning}\label{data-source-and-cleaning}}

The data comes from death registrations which are entered into the
Registry Services Tracking System (RSTS) by Registry Services staff who
are located at three of the civic centres, Etobicoke, North York and
Scarborough (Open Data Toronto 2023). It's creation supports the Vital
Statistics Act, a Provincial legislation (Open Data Toronto 2023) which
involve the collection of deaths, marriages, stillbirths, and live
births (Statistics Canada 2022b). This data set is updated monthly (Open
Data Toronto 2023). The variables from this data set represent the civic
centres; ``Etobicoke'', ``North York'', and ``Scarborough'', number of
death licenses registered in the month, place of death; ``Outside City
Limits'' or ``Toronto'', and time period by month and year in which the
death was registered (Open Data Toronto 2023).

To properly analyze this data for the purposes of this paper certain
data was removed including, all data for the years 2011-2015 and
2021-2023. This was done to ensure that the data set being worked with
had sufficient and consistent data to draw from without being too broad
in scope. Additionally, rows from the civic centre variable labelled
``Toronto'' were removed as death registrations at this location were
especially low and its data input, sporadic. Finally, the variable for
place of death was not included in the analysis of this data as it was
not relevant to its narrative.

\hypertarget{data-analysis}{%
\subsection{2.3 Data Analysis}\label{data-analysis}}

The province of Ontario represents a large portion of deaths in Canada,
taking on an estimated 37\% of Canadian mortality from 2016-2020
(Statistics Canada n.d.). It is also estimated that on average, that
just over 100 thousand people have died each year in Ontario between
2016 and 2020 (Statistics Canada n.d.). We may suggest then, that
Toronto makes up a large portion of this mortality rate, as it makes up
approximately 44\% of the Ontarian population (Statistics Canada n.d.).
Additionally, Canada has an aging population, meaning that it's
mortality rate is expected to rise every year although it rose
exponentially due to the COVID-19 pandemic (Statistics Canada 2022a).
Therefore, we may suspect that the COVID-19 pandemic would increase the
mortality rate significantly in Toronto as Canada saw an increase of
7.7\% in deaths in 2020 (Statistics Canada 2022a).

\begin{figure}

\caption{\label{fig-1}Death Licenses Provided by All Civic Centres
Yearly From 2016-2020}

{\centering \includegraphics{paper_files/figure-pdf/fig-1-1.pdf}

}

\end{figure}

When looking at Figure~\ref{fig-1}, however, we can see that from 2016
to 2018 to there was an increase in death licenses registered, as
expected, followed by a slight drop in 2019. This drop in 2019 may be
indicative of a backlog from Ontario's Vital Statistics in early 2019
which was cited as having delays of over three months, waiting for
death, birth and marriage certificates (Jeffords 2019). What is most
important in Figure~\ref{fig-1} is that when looking at the ``2020'' bar
in Figure~\ref{fig-1}, we can see that the death licenses registered
drops significantly in comparison to any of the previous 4 months.

\begin{figure}

\caption{\label{fig-2}Death Licenses Provided by Month from 2016-2020 at
Three Toronto Civic Centres}

{\centering \includegraphics{paper_files/figure-pdf/fig-2-1.pdf}

}

\end{figure}

when looking at Figure~\ref{fig-2} we can see - death licenses provided
by month and year from all three civic centres - highlights the spike in
death licenses provided in May - this may indicate a spike in deaths
near the beginning of COVID-19 pandemic, - \textasciitilde need to find
source to make this make sense\textasciitilde{}

\begin{figure}

\caption{\label{fig-3}Death Licenses Provided by Each Civic Centre
Yearly and Monthly from 2016-2020}

{\centering \includegraphics{paper_files/figure-pdf/fig-3-1.pdf}

}

\end{figure}

Looking briefly at Figure~\ref{fig-3} we can see a surprising drop in
death licenses provided in 2020. \#\#\#\#\#\# this is surprising as
death rates are said to have increased during the pandemic, yet there is
a signficant decline \#\#\#\#\#\# why? explain\ldots{}

\hypertarget{tbl-1}{}
\begin{table}
\caption{\label{tbl-1}Average Number of Deaths Per Month from 2016-2020 at Etobicoke Centre }\tabularnewline

\centering
\begin{tabular}{lrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{5}{c}{Average Number of Deaths} \\
\cmidrule(l{3pt}r{3pt}){2-6}
Month & 2016 & 2017 & 2018 & 2019 & 2020\\
\midrule
Jan & 342 & 360 & 372 & 346 & 358\\
Feb & 264 & 276 & 275 & 259 & 250\\
Mar & 275 & 281 & 285 & 269 & 237\\
Apr & 236 & 209 & 211 & 237 & 273\\
May & 315 & 320 & 329 & 320 & 403\\
Jun & 236 & 238 & 235 & 253 & 196\\
Jul & 223 & 250 & 263 & 275 & 235\\
Aug & 241 & 239 & 211 & 230 & 210\\
Sep & 248 & 225 & 251 & 236 & 185\\
Oct & 281 & 275 & 275 & 278 & 226\\
Nov & 348 & 351 & 340 & 356 & 280\\
Dec & 229 & 209 & 233 & 228 & 217\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{tbl-2}{}
\begin{table}
\caption{\label{tbl-2}Average Number of Deaths Per Month from 2016-2020 at Scarborough Centre }\tabularnewline

\centering
\begin{tabular}{lrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{5}{c}{Average Number of Deaths} \\
\cmidrule(l{3pt}r{3pt}){2-6}
Month & 2016 & 2017 & 2018 & 2019 & 2020\\
\midrule
Jan & 350 & 366 & 376 & 371 & 359\\
Feb & 266 & 269 & 272 & 289 & 262\\
Mar & 275 & 281 & 273 & 284 & 244\\
Apr & 273 & 262 & 261 & 268 & 299\\
May & 382 & 396 & 409 & 382 & 419\\
Jun & 214 & 214 & 201 & 206 & 223\\
Jul & 253 & 248 & 263 & 253 & 274\\
Aug & 275 & 279 & 252 & 271 & 288\\
Sep & 252 & 243 & 244 & 236 & 195\\
Oct & 256 & 264 & 258 & 244 & 226\\
Nov & 246 & 256 & 265 & 261 & 301\\
Dec & 214 & 184 & 199 & 203 & 272\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{tbl-3}{}
\begin{table}
\caption{\label{tbl-3}Average Number of Deaths Per Month from 2016-2020 at North York Centre }\tabularnewline

\centering
\begin{tabular}{lrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{5}{c}{Average Number of Deaths} \\
\cmidrule(l{3pt}r{3pt}){2-6}
Month & 2016 & 2017 & 2018 & 2019 & 2020\\
\midrule
Jan & 291 & 323 & 330 & 303 & 293\\
Feb & 210 & 214 & 236 & 222 & 193\\
Mar & 224 & 221 & 235 & 225 & 182\\
Apr & 184 & 164 & 155 & 180 & 189\\
May & 246 & 250 & 259 & 268 & 372\\
Jun & 194 & 193 & 199 & 179 & 180\\
Jul & 176 & 186 & 194 & 211 & 209\\
Aug & 199 & 195 & 165 & 172 & 186\\
Sep & 198 & 181 & 194 & 170 & 135\\
Oct & 183 & 192 & 205 & 176 & 160\\
Nov & 194 & 200 & 204 & 220 & 233\\
Dec & 165 & 139 & 151 & 184 & 189\\
\bottomrule
\end{tabular}
\end{table}

\clearpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-citejanitor}{}}%
Firke, Sam. 2021. \emph{Janitor: Simple Tools for Examining and Cleaning
Dirty Data}. \url{https://CRAN.R-project.org/package=janitor}.

\leavevmode\vadjust pre{\hypertarget{ref-citeopendatatoronto}{}}%
Gelfand, Sharla. 2022. \emph{Opendatatoronto: Access the City of Toronto
Open Data Portal}.
\url{https://CRAN.R-project.org/package=opendatatoronto}.

\leavevmode\vadjust pre{\hypertarget{ref-citectvnews}{}}%
Jeffords, Shawn. 2019. {``Delays in Birth, Death, Marriage Certificates
Caused by High Demand, Online Issues.''} CTV News Windsor.
\url{https://windsor.ctvnews.ca/delays-in-birth-death-marriage-certificates-caused-by-high-demand-online-issues-1.4305301}.

\leavevmode\vadjust pre{\hypertarget{ref-citehere}{}}%
Müller, Kirill. 2020. \emph{Here: A Simpler Way to Find Your Files}.
\url{https://CRAN.R-project.org/package=here}.

\leavevmode\vadjust pre{\hypertarget{ref-citercolorbrewer}{}}%
Neuwirth, Erich. 2022. \emph{RColorBrewer: ColorBrewer Palettes}.
\url{https://CRAN.R-project.org/package=RColorBrewer}.

\leavevmode\vadjust pre{\hypertarget{ref-citeopendatatoronto2}{}}%
Open Data Toronto. 2023. {``Death Registry Statistics.''} City of
Toronto Open Data Portal.
\url{https://open.toronto.ca/dataset/death-registry-statistics/}.

\leavevmode\vadjust pre{\hypertarget{ref-citepatchwork}{}}%
Pedersen, Thomas Lin. 2022. \emph{Patchwork: The Composer of Plots}.
\url{https://CRAN.R-project.org/package=patchwork}.

\leavevmode\vadjust pre{\hypertarget{ref-citeR}{}}%
R Core Team. 2020. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-citestatscan1}{}}%
Statistics Canada. n.d. {``Table 17-10-0008-01 Estimates of the
Components of Demographic Growth, Annual.''}
\url{https://doi.org/10.25318/1710000801-eng}.

\leavevmode\vadjust pre{\hypertarget{ref-citestatscan2}{}}%
---------. n.d. {``Table 17-10-0135-01 Population Estimates, July 1, by
Census Metropolitan Area and Census Agglomeration, 2016 Boundaries.''}
\url{https://doi.org/10.25318/1710013501-eng}.

\leavevmode\vadjust pre{\hypertarget{ref-citestatscan5}{}}%
---------. 2021. {``Provisional Death Counts and Excess Mortality,
January 2020 to August 2021.''}
\url{https://www150.statcan.gc.ca/n1/daily-quotidien/211108/dq211108a-eng.htm}.

\leavevmode\vadjust pre{\hypertarget{ref-citestatscan4}{}}%
---------. 2022a. {``Deaths, 2020.''}
\url{https://www150.statcan.gc.ca/n1/daily-quotidien/220124/dq220124a-eng.htm}.

\leavevmode\vadjust pre{\hypertarget{ref-citestatscan3}{}}%
---------. 2022b. {``Frequently Asked Questions on Vital Statistics.''}
\url{https://www.statcan.gc.ca/en/about/relevant/vscc/faq}.

\leavevmode\vadjust pre{\hypertarget{ref-citeggplot2}{}}%
Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
Analysis}. Springer-Verlag New York.
\url{https://ggplot2.tidyverse.org}.

\leavevmode\vadjust pre{\hypertarget{ref-citetidyverse}{}}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\leavevmode\vadjust pre{\hypertarget{ref-citedplyr}{}}%
Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2022.
\emph{Dplyr: A Grammar of Data Manipulation}.
\url{https://CRAN.R-project.org/package=dplyr}.

\leavevmode\vadjust pre{\hypertarget{ref-citeknitr}{}}%
Xie, Yihui. 2023. \emph{Knitr: A General-Purpose Package for Dynamic
Report Generation in r}. \url{https://yihui.org/knitr/}.

\leavevmode\vadjust pre{\hypertarget{ref-citekablextra}{}}%
Zhu, Hao. 2021. \emph{kableExtra: Construct Complex Table with 'Kable'
and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.

\end{CSLReferences}

\end{document}

### 2023-02-03 scripts/03-data_cleaning.R

@@ -14,7 +14,6 @@
 #### install.packages("knitr") ####
 #### install.packages("RColorBrewer") ####
 #### install.packages("kableExtra") ####
-#### install.packages("bookdown") ####
 #### install.packages("rmarkdown") ####
 #### install.packages("patchwork") ####
 #### install.packages("here") ####
@@ -27,7 +26,6 @@ library(dplyr)
 library(knitr)
 library(RColorBrewer)
 library(kableExtra)
-library(bookdown)
 library(patchwork)
 library(readr)
 library(here)
@@ -84,15 +82,15 @@ cleaned_death_registry_data <- cleaned_death_registry_data |>
       ) 
   )
 
-#### Ordering Months #####
-cleaned_death_registry_data$Month <- factor(cleaned_death_registry_data$Month, levels=c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
-
 #### Filtering Out "Toronto" Civic Centre from Data Due to Missing Data ####
 cleaned_death_registry_data <- cleaned_death_registry_data |>
   filter(!grepl('TO', civic_centre))
 
-#### turning years into factor ####
+#### Turning Years Into Factor ####
 cleaned_death_registry_data$Year <- as.factor(cleaned_death_registry_data$Year)
+
+#### Ordering Months #####
+cleaned_death_registry_data$Month <- factor(cleaned_death_registry_data$Month, levels=c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
 
 #### Creating a CSV File for Cleaned Data ####
 write_csv(

### 2023-02-03 scripts/04-tbl_data_cleaning_etobicoke.R

@@ -1,0 +1,88 @@
+#### Preamble ####
+# Purpose: Read in data from Open Data Toronto and clean the civic centre data from 2012 downloaded from Open Data Toronto's statistics on death registration.Link: https://open.toronto.ca/dataset/death-registry-statistics/
+# Author: Chloe Thierstein
+# Email: chloe.thierstein@utoronto.ca
+# Date: 2 February 2023
+# Prerequisites: Need to know where to find the City of Toronto death registry data
+
+#### Creating Datasets for 5 Year Period by Month at Etobicoke Civic Centre ####
+
+#### Repeat Process for Each Year (2016-2020) ####
+
+#### 2016 ####
+et_avg_num_deaths_by_month_2016 <-cleaned_death_registry_data |>
+  filter(Year == "2016" | civic_centre == "Etobicoke") |> #filter for year and civic centre
+  group_by(Month) |> #group by month
+  summarize("Average # of Deaths Registered" = mean(death_licenses)) #find average of death licenses given out 
+
+#### 2017 ####
+et_avg_num_deaths_by_month_2017 <-cleaned_death_registry_data |>
+  filter(Year == "2017" | civic_centre == "Etobicoke") |>
+  group_by(Month) |> 
+  summarize("Average # of Deaths Registered" = mean(death_licenses)) 
+
+#### 2018 ####
+et_avg_num_deaths_by_month_2018 <-cleaned_death_registry_data |>
+  filter(Year == "2018" | civic_centre == "Etobicoke") |>
+  group_by(Month) |> 
+  summarize("Average # of Deaths Registered" = mean(death_licenses)) 
+
+#### 2019 ####
+et_avg_num_deaths_by_month_2019 <-cleaned_death_registry_data |>
+  filter(Year == "2019" | civic_centre == "Etobicoke") |>
+  group_by(Month) |> 
+  summarize("Average # of Deaths Registered" = mean(death_licenses)) 
+
+#### 2020 ####
+et_avg_num_deaths_by_month_2020 <-cleaned_death_registry_data|>
+  filter(Year == "2020" | civic_centre == "Etobicoke") |>
+  group_by(Month) |> 
+  summarize("Average # of Deaths Registered" = mean(death_licenses))
+
+
+
+#### Combining Datasets for 5 Year Period by Month at Etobicoke Civic Centre ####
+
[…]

### 2023-02-03 outputs/paper/paper.qmd
updated readme and license

@@
 ---
-title: "Death Registrations May Indicate a Discreptency Between Expected and Reported Data During the COVID-19 Pandemic"
-subtitle: "An Analysis of Death Registrations from 2016-2020"
+title: "Death Registrations May Indicate a Discreptency in Reported Data During the COVID-19 Pandemic"
+subtitle: "An Analysis of Death Registrations in Toronto from 2016-2020"
 author: "Chloe Thierstein"
@@
   theme(legend.position = "none")+
-  scale_fill_brewer(palette="Set3")
+  scale_fill_brewer(palette="Set2")
 
@@
   ) +
-  scale_fill_brewer(palette = "Set3")
+  scale_fill_brewer(palette = "Set2")
 ```

### 2023-02-03 scripts/04-tbl_data_cleaning_etobicoke.R

@@ -1,88 +1,0 @@
-#### Preamble ####
-# Purpose: Read in data from Open Data Toronto and clean the civic centre data from 2012 downloaded from Open Data Toronto's statistics on death registration.Link: https://open.toronto.ca/dataset/death-registry-statistics/
-# Author: Chloe Thierstein
-# Email: chloe.thierstein@utoronto.ca
-# Date: 2 February 2023
-# Prerequisites: Need to know where to find the City of Toronto death registry data
-
-#### Creating Datasets for 5 Year Period by Month at Etobicoke Civic Centre ####
-
-#### Repeat Process for Each Year (2016-2020) ####
-
-#### 2016 ####
-et_avg_num_deaths_by_month_2016 <-cleaned_death_registry_data |>
-  filter(Year == "2016" | civic_centre == "Etobicoke") |> #filter for year and civic centre
-  group_by(Month) |> #group by month
-  summarize("Average # of Deaths Registered" = mean(death_licenses)) #find average of death licenses given out 
-
-#### 2017 ####
-et_avg_num_deaths_by_month_2017 <-cleaned_death_registry_data |>
-  filter(Year == "2017" | civic_centre == "Etobicoke") |>
-  group_by(Month) |> 
-  summarize("Average # of Deaths Registered" = mean(death_licenses)) 
-
-#### 2018 ####
-et_avg_num_deaths_by_month_2018 <-cleaned_death_registry_data |>
-  filter(Year == "2018" | civic_centre == "Etobicoke") |>
-  group_by(Month) |> 
-  summarize("Average # of Deaths Registered" = mean(death_licenses)) 
-
-#### 2019 ####
-et_avg_num_deaths_by_month_2019 <-cleaned_death_registry_data |>
-  filter(Year == "2019" | civic_centre == "Etobicoke") |>
-  group_by(Month) |> 
-  summarize("Average # of Deaths Registered" = mean(death_licenses)) 
-
-#### 2020 ####
-et_avg_num_deaths_by_month_2020 <-cleaned_death_registry_data|>
-  filter(Year == "2020" | civic_centre == "Etobicoke") |>
-  group_by(Month) |> 
-  summarize("Average # of Deaths Registered" = mean(death_licenses))
-
-
-
-#### Combining Datasets for 5 Year Period by Month at Etobicoke Civic Centre ####
-
[…]

### 2023-02-03 outputs/paper/paper.qmd
continue writing

@@
 #| label: fig-1
-#| fig-cap: "Death Licenses Provided by All Civic Centres Yearly From 2016-2020 "
+#| fig-cap: "Death Licenses Registered at Three Toronto Civic Centres from 2016-2020"
 
@@
 What is most important in @fig-1 is that when looking at the "2020" bar in @fig-1, we can see that the death licenses registered drops significantly in comparison to any of the previous 4 months.
+However, the amount of death licenses registered in 2020 was inconsistent as can be seen in @fig-2.
 
@@
 #| label: fig-2
-#| fig-cap: "Death Licenses Provided by Month from 2016-2020 at Three Toronto Civic Centres"
+#| fig-cap: "Death Licenses Provided by Month at Three Toronto Civic Centres from 2016-2020"
 

### 2023-02-04 outputs/paper/paper.qmd
Analysis added

@@
 ---
-title: "Death Registrations May Indicate a Discreptency in Reported Data During the COVID-19 Pandemic"
-subtitle: "An Analysis of Death Registrations in Toronto from 2016-2020"
+title: "Death License Registrations May Indicate a Discreptency in Reported Data During the COVID-19 Pandemic"
+subtitle: "An Analysis of Death License Registrations in Toronto from 2016-2020"
 author: "Chloe Thierstein"
 thanks: "Code and data are available at: https://github.com/cthierst/death_registry_analysis.git."
-date: "`r Sys.time()`"
+date: "February 3, 2023"
 date-format: "D MMMM YYYY"
-abstract: "The COVID-19 pandemic marked a significant shift in Canadian mortality rates in the year 2020.
-Using death registry data from three civic centres in the greater Toronto are from 2016 to 2020, we demonstrate that even data from a considerably reliable source may become distorted"
-format: pdf
+format: 
+  pdf:
+    documentclass: article
+    geometry: margin = 1in
+    abstract:  "Death registry statistics offer us a useful tool to better out healthcare system, from monitoring trends to government policy, it is essential to our understanding of the world around us.
+The COVID-19 pandemic marked a significant shift in the Canadian mortality rate in the year 2020.
+This paper utilizes death registry data from three civic centres in the greater Toronto area from 2016 to 2020, alongside secondary sources and research.
+From this work, we demonstrate that even data from considerably reliable sources may become distorted if enough care is not taken in its recording and the world around the data is not considered."
 bibliography: references.bib
@@
 We reason that this discrepancy in expected data versus reported data could be the result of strain and complications put onto record-keeping during this time similarly to what was seen in early 2019 with long delays in the distribution of death, birth and marriage certificates due to high online demand [@citectvnews].
-Future work could look specifically at how strains on data reporting centres during the COVID-19 pandemic influenced data.
+Future work could look specifically at how strain on data reporting centres during the COVID-19 pandemic influenced data and mitigation techniques for the future
 
@@
 Next, we discuss in more detail, death licenses provided by each civic centre to determine a possible cause for the uncharacteristic decline of death licenses in 2020.
-We will then compare the mean number of deaths from each civic centre to further gauge how these centres relate to rates of death registrations.
+We will then look at the mean number of deaths from each civic centre to further gauge how these centres relate to rates of death registrations.
 Finally, we will consider the limitations of the data, including biases.
@@
 #### Import cleaned_death_registry.csv ####
-cleaned_death_registry_data <- read.csv(here::here("/Users/chloethierstein/Desktop/death_registry_analysis/inputs/data/cleaned_death_registry.csv"))
+cleaned_death_registry_data <- read.csv(here::here("./inputs/data/cleaned_death_registry.csv"))
 
@@
     x = "Year",
-    y = "Death Licenses"
+    y = "# of Death Licenses"
   ) +
@@
 What is most important in @fig-1 is that when looking at the "2020" bar in @fig-1, we can see that the death licenses registered drops significantly in comparison to any of the previous 4 months.
-However, the amount of death licenses registered in 2020 was inconsistent as can be seen in @fig-2.
+However, the amount of death licenses registered each month in 2020 was inconsistent as can be seen in @fig-2.
 
@@
 #| label: fig-2
-#| fig-cap: "Death Licenses Provided by Month at Three Toronto Civic Centres from 2016-2020"
+#| fig-cap: "Death Licenses Registered by Month at Three Toronto Civic Centres from 2016-2020"
 
@@
     x = "Month",
-    y = "Death Licenses"
+    y = "# of Death Licenses"
   ) +
@@
 
-when looking at @fig-2 we can see - death licenses provided by month and year from all three civic centres - highlights the spike in death licenses provided in May - this may indicate a spike in deaths near the beginning of COVID-19 pandemic, - \~need to find source to make this make sense\~
+\newpage
+
+By looking at @fig-2 we can see that, generally, each month is relatively consistent in the amount of death licenses registered from the three civic centres from 2016 to 2019 with a few variations.
+However, when looking at 2020, we notice significantly lower rates of death license registration each month overall in comparison to past years with the exception of May and November.
+In the later, we see death license registrations that are nearly balanced with previous years.
+Whereas in May, there is a large spike in death license registration that far surpasses any month's death license registration in the previous four years.
+
+Looking at May's spike without knowing the context of the years, we may assume that it is a particularly deadly month, however that we can see that the spike in May, shown in @fig-2, is likely indicative of the first wave of COVID-19 in Canada hitting its peak at the beginning of the month before declining [@citejamanetwork].
+This makes sense, however, COVID-19 infections were not the only factor driving the increased death rates in Canada during 2020.
+There were indirect consequences of the pandemic which lead to the higher death rate in Canada.
+For example, there were over four-thousand deaths from accidental poisonings which was nearly one-thousand more than in 2019 [@citestatscan4].
+By understanding that the increase in the Canadian death rate in 2020 was not only caused by COVID-19 infections, we can see the complexity of the situation.
+Although the continued decline in death licenses registered from these civic centres during a year that saw an unexpected 5.2% [@citestatscan5] rise in deaths indicates a potential issue in the reporting done at these civic centres.
+
+@fig-3 demonstrates how many death licenses were registered at each civic centre, Etobicoke, North York and Scarborough from 2016 to 2020, both monthly and yearly.
+From @fig-3 we can see that, generally, Scarborough registers the most death licenses, followed by Etobicoke and finally North York.
+Interestingly, in 2020, Etobicoke largely overtook Scarborough in death license registrations while simulataneously registered more death licenses than it had in the previous four years.
+This could indicate that the Etobicoke centre was taking on more work from other civic centres as it was perhaps less overwhelmed during the COVID-19 pandemic.
+However, more data would be needed to confirm this.
 
@@
     x = "Month",
-    y = "Death Licenses",
+    y = "# Death Licenses",
     fill = "Civic Centre"
@@
     x = "Year",
-    y = "Death Licenses",
+    y = "# Death Licenses",
     fill = "Civic Centre"
@@
 #| label: fig-3
-#| fig-cap: "Death Licenses Provided by Each Civic Centre Yearly and Monthly from 2016-2020"
+#| fig-cap: "Monthly and Yearly Number of Death Licenses Registered by Each Civic Centre from 2016-2020"
 
@@
 ```
-
-Looking briefly at @fig-3 we can see a surprising drop in death licenses provided in 2020.
-\###### this is surprising as death rates are said to have increased during the pandemic, yet there is a signficant decline \###### why?
-explain...
 
@@
   ) |>
-  add_header_above(c(" " = 1, "Average Number of Deaths" = 5))
+  add_header_above(c(" " = 1, "# of Death License Registrations" = 5))
 
@@
     linesep = "") |>
-  add_header_above(c(" " = 1, "Average Number of Deaths" = 5))
+  add_header_above(c(" " = 1, "# of Death License Registrations" = 5))
 
@@
     linesep = "") |>
-  add_header_above(c(" " = 1, "Average Number of Deaths" = 5)) 
+  add_header_above(c(" " = 1, "# of Death License Registrations" = 5)) 
 ```
+
+\newpage
+
+Many of the trends spoken about previously can be seen explicitly in @tbl-1, @tbl-2, and @tbl-3.
+These tables present the average number of death licenses registered per month from 2016 to 2020 at each civic centre.
+These tables allow us to better grasp the variability in death licenses registered each month from 2016-2020.
+High variations in amount of death licenses registered can be seen mostly between months, although between years these variations become much smaller.
+This is important as upon first glance one would not assume inherently that the data actually shows much less death license registrations in 2020.
+
+While more data and research needs to be done to fully understand the story of data collection and reporting during the COVID-19 pandemic, this paper looks at a screenshot of death registry data from three civic centres between 2016 and 2020.
+By doing this, we attempt to better understanding the significance of data collection and how it may be impacted by outside conditions.
+We suggest that there is data missing from the report due to a variety of factors and reason that pressure and complications but onto data-keeping centres during the COVID-19 pandemic may be to blame.
+We encourage future inquiries to look specifically at how strain put on data reporting centres during the COVID-19 pandemic influenced data and how to mitigate these strains in the future.
 

### 2023-02-04 scripts/00-simulation.R
Combine simulation and tests together

@@ -1,11 +1,10 @@
 #### Preamble ####
-# Purpose: Simulate from Open Data Toronto's statistics on death registration. Link: https://open.toronto.ca/dataset/death-registry-statistics/
+# Purpose: Simulate and test data based on Open Data Toronto's statistics on death registration. Link: https://open.toronto.ca/dataset/death-registry-statistics/
 # include columns ID, Death Licenses, Civic Centre, Place of Death, Month (as placeholder for date)
 # Author: Chloe Thierstein
 # Email: chloe.thierstein@mail.utoronto.ca
 # Date: 23 January 2023
 # Prerequisites: Need to know where to find the City of Toronto death registry data
-
 
 
 library(tibble)

### 2023-02-04 scripts/01-simulation_tests.R
Remove separate tests script

@@ -1,23 +1,0 @@
-#### Preamble ####
-# Purpose: Test data from Open Data Toronto's statistics on death registration. Link: https://open.toronto.ca/dataset/death-registry-statistics/
-# include columns ID, Death Licenses, Civic Centre, Place of Death, Month (as placeholder for date)
-# Author: Chloe Thierstein
-# Email: chloe.thierstein@mail.utoronto.ca
-# Date: 23 January 2023
-# Prerequisites: Need to know where to find the City of Toronto death registry data
-
-#### testing simulated data ####
-
-#### test 1 #### 
-
-simulated_data$"Civic Centre" |>
-  unique() |>
-  length() == 3
-
-#### test 2 #### 
-
-simulated_data$"Year" |> min() == 2016
-
-#### test 3 ####
-
-simulated_data$"Death Licenses" |> class() == "numeric"

### 2023-02-04 scripts/03-data_cleaning.R
Update note on packages used

@@ -8,16 +8,19 @@
 
 #### Workspace Set-up ####
 
-#### install.packages(tidyverse) ####
-#### install.packages(janitor) ####
-#### install.packages(ggplot2) ####
+#### Need to Install These Packages to Run Script ####
+#### install.packages("tidyverse") ####
+#### install.packages("opendatatoronto") ####
+#### install.packages("janitor") ####
+#### install.packages("ggplot2") ####
 #### install.packages("knitr") ####
 #### install.packages("RColorBrewer") ####
 #### install.packages("kableExtra") ####
-#### install.packages("rmarkdown") ####
 #### install.packages("patchwork") ####
+#### install.packages("readr") ####
 #### install.packages("here") ####
 
+#### Loading in Libraries ####
 library(tidyverse)
 library(opendatatoronto)
 library(janitor)

### 2023-02-04 outputs/paper/paper.qmd
Updated title

@@
 ---
-title: "Death License Registrations May Indicate a Discreptency in Reported Data During the COVID-19 Pandemic"
+title: "Death License Registrations May Indicate a Discrepency in Reported Data During the COVID-19 Pandemic"
 subtitle: "An Analysis of Death License Registrations in Toronto from 2016-2020"

### 2023-02-04 scripts/03-data_cleaning.R
Update Preamble

@@ -1,6 +1,6 @@
 
 #### Preamble ####
-# Purpose: Read in data from Open Data Toronto and clean the civic centre data from 2012 downloaded from Open Data Toronto's statistics on death registration.Link: https://open.toronto.ca/dataset/death-registry-statistics/
+# Purpose: Clean data from Open Data Toronto and clean the civic centre data from 2012 downloaded from Open Data Toronto's statistics on death registration.Link: https://open.toronto.ca/dataset/death-registry-statistics/
 # Author: Chloe Thierstein
 # Email: chloe.thierstein@utoronto.ca
 # Date: 23 January 2023

### 2023-02-04 scripts/02-download_data.R
Update workspace set-up

@@ -5,6 +5,16 @@
 # Date: 23 January 2023
 # Prerequisites: Need to know where to find the City of Toronto death registry data
 
+#### Workspace Set-up ####
+
+#### Installing Packages ####
+
+#### install.packages("tidyverse")
+
+#### Opening Libraries ####
+
+library(tidyverse)
+
 #### Read in Data ####
 
 raw_death_registry_data <-
